---
title: "DC-analysis-new"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


PTG (ME49)  = type II (considered to be avirulent) with high expression/levels of IL-12
LDM (RH/GT1) = type I (considered virulent) 

```{r, load required packages, warning=FALSE}

library(Seurat)
library(dplyr)
library(Matrix)
library(gtable)
library(grid)
library(gridExtra)
library(rlang)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(ggplot2)
library(biomaRt)
library(tidyr)
library(plotly)
library(pheatmap)
library(forcats)
library(patchwork)
library(sctransform)
library(openxlsx)
suppressMessages(require(biomaRt))
library(EnhancedVolcano)
library(pheatmap)
library(RCurl)
library(DESeq2)
library(WGCNA)
library(ggridges)
library(clustifyr)
library(clusterProfiler)
library(org.Mm.eg.db)
library(enrichplot)

```

```{r, define colors for paper}

grayscale = c("gray5","gray15", "gray30", "gray55", "gray70", "gray85", "gray95", "gray100")

#toxo color scale
toxo.c = c("darkgreen", "olivedrab3", "maroon4", "hotpink2")


#cluster colors
c.cols <- c("#AD1457", "#AB47BC", "#827717", "#2E7D32","#0097A7", "#E64A19", "blue2", "#7E57C2", "#BF360C")

#cellclass colors
cell.cols <- c("dodgerblue2", "darkgreen", "olivedrab3", "palegoldenrod","red4", "maroon4", "hotpink2","rosybrown1")

#celltype colors 

type.cols <- c("turquoise4", "mediumorchid4")

m.cycle.c <- c("darkolivegreen", "darkorange3", "royalblue4")

```



```{r, import data from mapping}

cnt_131 <- read.table("~/t.gondi_dc/data/cnt/feat_count/count131.csv.gz", header = T,sep = "\t", row.names = 1)
head(cnt_131)
cnt_132 <- read.table("~/t.gondi_dc/data/cnt/feat_count/count132.csv.gz", header = T,sep = "\t", row.names = 1)

meta <- read.table("~/t.gondi_dc/data/meta/properanno.csv", header = T, sep = ",", row.names = 1)
head(meta)

meta2 <- read.table("~/t.gondi_dc/data/meta/latest_mapping.csv", header = T, sep = ",")

meta_131 <- meta2[grep(pattern = "131_", meta$cellname),]
meta_132 <- meta2[grep(pattern = "132_", meta$cellname),]

```

```{r, format cnt_marix}

#plate1
cnt_131 <- cnt_131[,6:389]
head(cnt_131)

names <- sapply(colnames(cnt_131),
               function(x) {strsplit(x,'_')[[1]][2]})
colnames(cnt_131) <- names
colnames(cnt_131) <- gsub("X", "", colnames(cnt_131))
colnames(cnt_131) <- gsub("131.", "", colnames(cnt_131))
colnames(cnt_131) <- paste0("131", sep = "_", colnames(cnt_131))
head(cnt_131)

#plate2
cnt_132 <- cnt_132[,6:389]
head(cnt_132)

names <- sapply(colnames(cnt_132),
               function(x) {strsplit(x,'_')[[1]][1]})
colnames(cnt_132) <- names
colnames(cnt_132) <- gsub("X", "", colnames(cnt_132))
colnames(cnt_132) <- gsub("132.", "", colnames(cnt_132))
colnames(cnt_132) <- paste0("132", sep = "_", colnames(cnt_132))
head(cnt_132)

dc.data <- cbind(cnt_131, cnt_132)

```



Import the count matrix of the single cell sequencing experiment and subset it, so it only contains the counts for the mouse genome

```{r, import expression data and subset it}

#import expression matrix first 

dc.data <- read.csv("~/t.gondi_dc/data/cnt/cnt_mus_toxo_raw.csv", row.names = 1, header = T, check.names = FALSE)

#subset for only mouse genes: 
mus.genes <- rownames(dc.data)[grep("ENSMUSG",rownames(dc.data))]

head(mus.genes)
dc.data.mus <- dc.data[mus.genes,]

```

```{r, get count for toxo genes}

toxo.genes <- rownames(dc.data)[grep("TGME49",rownames(dc.data))]
dc.data.toxo <- dc.data[toxo.genes,]

dc.data.toxo[1:5,1:5]
nUMI_toxo <-  as.data.frame(colSums(dc.data.toxo))

se.dc.mus <- AddMetaData(se.dc.mus, metadata = nUMI_toxo, col.name = "nUMI_toxo")


```

Convert ensembl-Ids to gene symbols in the matrix

```{r make gene symbol annotation file}
#Create a vector from the column names 
gn <- as.vector(rownames(dc.data.mus))
gn <- sapply(gn, function(x) {strsplit(x,'\\.')[[1]][1]} )
gn <- as.vector(unlist(gn))
rownames(dc.data.mus) <- gn

#Create an annotation file 

#mouse
mus_ensembl <- AnnotationDbi::mapIds(org.Mm.eg.db::org.Mm.eg.db,
                                      keys=gn,
                                      column="SYMBOL",
                                      keytype="ENSEMBL",
                                      multiVals="first")

#plasmodium



#Select only the annotations that are not NA 
keep <- (!(is.na(mus_ensembl)))
genes <- mus_ensembl[keep]

#get the names only
extract <- names(mus_ensembl)[keep] 
##remove duplicates
#For ensIds
extract <- as.vector(extract[!(duplicated(extract))])
#For gene symbols
genes <- as.vector(genes[!(duplicated(extract))])

#for the expression matrix
dc.data.mus <- dc.data.mus[extract,]

#give new column names with gene symbols
rownames(dc.data.mus) <- make.unique(genes)

nmat[1:5,1:5]

```

Create the seurat object for analysis - set 2 markers for "mature" and "immature" cell types of BDMCs to get an impression of the proportions of these two types for each condition in the data-set. However I would like to use the mature/immature classification we have calculated - find a way to use this instead if comparing numbers 

```{r, Create Seuratobject, width = 10, height = 40}

meta1 <- read.table("~/t.gondi_dc/data/meta/metadata-200430.csv", sep = ",", header = T, row.names = 1)
#meta_new <- read.table("~/t.gondi_dc/meta_141221.csv",sep = ";", header = T, row.names = 1)

se.dc.mus <- CreateSeuratObject(nmat, min.cells = 3, min.features = 200, project = "DC-TG-FH-M", meta.data = meta_new)

se.dc.mus[["percent.mt"]] <- PercentageFeatureSet(se.dc.mus,pattern = "mt")
se.dc.mus[["percent.mature"]] <- PercentageFeatureSet(se.dc.mus, features = "mature")
se.dc.mus[["percent.immature"]] <- PercentageFeatureSet(se.dc.mus, features= "immature")

grayscale = c("gray5","gray15", "gray30", "gray55", "gray70", "gray85", "gray95", "gray100")

VlnPlot(se.dc.mus, features = c("nCount_RNA","nFeature_RNA", "percent.mt", "nUMI_toxo"), ncol = 2, group.by = "cell_class", cols = grayscale)

pdf("~/t.gondi_dc/res/figures/fig1/vln_toxo.pdf", width = 15, height = 7.5)
VlnPlot(se.toxo, features = c("TGME49-259020","TGME49-293690", "TGME49-201780", "TGME49-233460"), group.by = "cell_class", cols = toxo.c , ncol = 2)
dev.off()

pdf("~/t.gondi_dc/res/figures/fig1/vln_toxo_fraction.pdf", width = 7.5, height = 7.5)
VlnPlot(se.toxo, features = "toxo_nUMI_p", group.by = "cell_class", cols = toxo.c)
dev.off()


pdf("~/t.gondi_dc/res/figures/fig1/vln_all.pdf", width = 15, height = 7.5)
VlnPlot(se.dc.mus, features = c("Rpl9", "Hprt", "Ifit1", "Egr1"), group.by = "cell_class", cols = cell.cols , ncol = 2)
dev.off()

#subset the cells with a high percentage of mitochondrial genes 
se.dc.mus <- subset(se.dc.mus, subset = percent.mt < 10)


#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/qc/RNA-andFeatureCount-DC.png", width = 35, height = 15, res = 300, units = "cm") #Units are pixels by default
print(VlnPlot(se.dc.mus, features = c("nCount_RNA", "nFeature_RNA"), ncol = 2, group.by = "cell_class", cols = grayscale))
#dev.off()

```

```{r, boxplot for number of toxo-transcripts / condition}

df <- data.frame(features = se.dc.mus$toxo_nUMI_p, condition = se.dc.mus$cell_class)
ggplot(df, aes(x= condition, y = features*100)) + 
  geom_boxplot()

VlnPlot(se.dc.mus, features = "toxo_nUMI", group.by = "seurat_clusters", cols = c.cols )

```


```{r, get data for cell specific transcript proportions}

genes.table.tg.mus <- read.delim("~/t.gondi_dc/data/meta/genes_annotation_TG_MUS.tab")

dc.data.p <- dc.data[intersect(rownames(dc.data), subset(genes.table.tg.mus, gene_biotype == "protein_coding")$ensembl_gene_id),]

#Toxoplasma and mouse specific genes stored in a dataframe

mouse_genes <- intersect(rownames(dc.data), subset(genes.table.tg.mus, species == "mouse")$ensembl_gene_id)

toxo_genes <- intersect(rownames(dc.data.p), subset(genes.table.tg.mus, species == "toxoplasmagondii")$ensembl_gene_id)

qc <- data.frame(nUMI = colSums(dc.data), 
                 mouse_nUMI = colSums(dc.data[mouse_genes, ]),
                 toxo_nUMI = colSums(dc.data[toxo_genes, ]),
                 nGene = apply(dc.data, 2, function(x) sum(x>2)),
                 mouse_nGene = apply(dc.data[mouse_genes, ], 2 , function(x) sum(x>2)),
                 toxo_nGene = apply(dc.data[toxo_genes, ], 2 , function(x) sum(x>2)),
                 cell_ID = factor(1:ncol(dc.data)), 
                 cell = colnames(dc.data))

qc[, c("mouse_nUMI_p", "toxo_nUMI_p")] <- qc[, c("mouse_nUMI", "toxo_nUMI")]/qc[, c("nUMI")]

qc[, c("mouse_nGene_p", "toxo_nGene_p")] <- qc[, c("mouse_nGene", "toxo_nGene")]/qc[, c("nGene")]

qc$cell_ID <- factor(qc$cell_ID, levels = order(qc$toxo_nUMI_p, decreasing = F))

options(repr.plot.width = 10, repr.plot.height = 3)
qcm <- reshape2::melt(qc, measure.vars = c("mouse_nUMI_p", "toxo_nUMI_p"))
qcm <- qcm[order(qcm$toxo_nUMI, decreasing = T), ]
qcm1 <- subset(qcm, qcm$cell %in% names(which(se.dc.mus$cell_class == "PTG_3h")))

pdf("~/t.gondi_dc/res/figures/fig1/jitter_nUMI_species.pdf", width = 5, height = 3)
    print(ggplot(qcm, aes(cell_ID, value, fill = ifelse(variable == "mouse_nUMI_p", "mouse", "toxo"), color = ifelse(variable == "mouse_nUMI_p", "mouse", "toxo_nUMI_p"))) +   geom_jitter() +
    scale_y_continuous(labels = scales::percent) + #"#4477AA", "#CC6677"
    theme(axis.text.x = element_blank()) +
    scale_fill_manual(values = c("mouse" = "#33658A", "toxo" = "#F26419")) +
    scale_color_manual(values = c("mouse" = "#33658A", "toxo" = "#F26419"), na.value = "#F26419")+ 
    ylab("amount of species specific transcripts [%]") + 
    xlab("cell") + 
    theme(legend.title = element_blank(), 
          panel.border = element_blank(),
          panel.background = element_rect("white"), 
          panel.grid.major.y = element_line(colour="lightgrey", size=0.5),
          panel.grid.major.x = element_blank()))
    dev.off()

```


```{r, normalize the data, results="hide"}

se.dc.mus <- SCTransform(se.dc.mus)

```

```{r, Run dimensionality reduction and clustering}

se.dc.mus <- RunPCA(se.dc.mus)


print(se.dc.mus[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(se.dc.mus, dims = 1:2, reduction = "pca")

DimPlot(se.dc.mus,  reduction = "pca", group.by = "cell_class", cols = "Dark2")

DimHeatmap(se.dc.mus, dims = 1:15, cells = 500, balanced = TRUE)

ElbowPlot(se.dc.mus, ndims= 20, reduction ="pca")

se.dc.mus <- FindNeighbors(se.dc.mus, dims = 1:15)
se.dc.mus <- FindClusters(se.dc.mus) #start with default resolution of 0.8
#inspect the clusters
head(Idents(se.dc.mus), 5)

se.dc.mus <- RunUMAP(se.dc.mus, dims = 1:10, reduction = "pca")
#min.dist let's me play around with the thightnes of the cells within the clusters
#the spread let's me play around with the coordinates of the distance between the clusters 
##default settings are -> min.dist: 0.3, spread: 1

#rename the clusters
new.cluster.ids <- c("M 1", "M 3", "M 4", "M 6", "M 7", "M 5", "M 2", "M 8","M 9")
names(new.cluster.ids) <- levels(se.dc.mus)
se.dc.mus <- RenameIdents(se.dc.mus, new.cluster.ids)
levels(se.dc.mus) <- as.factor(c("M 1", "M 2", "M 3","M 4", "M 5", "M 6", "M 7", "M 8", "M 9"))
levels(se.dc.mus)
       

#alternatively we can rename the seurat clusters colomun right away:
se.dc.mus$seurat_clusters <- ifelse(se.dc.mus$seurat_clusters == "1", paste("M 3"), ifelse(se.dc.mus$seurat_clusters == "0", paste("M 1"), ifelse(se.dc.mus$seurat_clusters == "6", paste("M 2"), ifelse(se.dc.mus$seurat_clusters == "4", paste("M 7"), ifelse(se.dc.mus$seurat_clusters == "2", paste("M 4"), ifelse(se.dc.mus$seurat_clusters == "5", paste("M 5"), ifelse(se.dc.mus$seurat_clusters == "3", paste("M 6"), ifelse(se.dc.mus$seurat_clusters == "7", paste("M 8"), paste("M 9")))))))))

DimPlot(se.dc.mus, group.by = "seurat_clusters", cols = c.cols)

#se.dc.mus <- readRDS("~/t.gondi_dc/res/seurat-objects/seurat_mus_only_210714.RDS")
```


Plot the Umap and visualize the clusters as well as the original cell identity

```{r, umap coordinates and plots}

umap.dc.m.clusters <- DimPlot(se.dc.mus, reduction = "umap",
                              cols = c.cols) +
                              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank())
umap.dc.m.cellclass <- DimPlot(se.dc.mus, reduction = "umap", 
                               group.by = "cell_class", 
                               cols = cell.cols)  +
                              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank(),plot.title = element_blank())

#extract the umap coordinates for possible cross-references
umap.coordinates.dc <- as.data.frame(Embeddings(se.dc.mus, reduction = "umap"))


pdf("~/t.gondi_dc/res/figures/fig1/dc_umap_clusters.pdf", width = 12, height = 5) #Units are pixels by default
print(cowplot::plot_grid(umap.dc.m.cellclass,umap.dc.m.clusters))
dev.off()


pdf("~/t.gondi_dc/res/figures/fig2/cell_types_map.pdf", width = 5, height = 4) #Units are pixels by default
DimPlot(se.dc.mus, reduction = "umap", 
                               group.by = "celltype", 
                               cols = type.cols)  +
                              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank())
dev.off()

umap.dc.m.clusters
umap.dc.m.cellclass
```


```{r, calculate percentages of each condition within individual clusters}

# sapply function to give vector for each cluster and condition and create a dataframe with the information 
clusters <- c(0:8)

LPS_perCluster <- sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LPS" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_3h_perCluster <- sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_12h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_12h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_12h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_12h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

ctrl_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "ctrl" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_Lys3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_Lys3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_Lys3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_Lys3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

n1 <- c("cluster0", "cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8")
identpercluster <- data.frame(n1, LDM_3h_perCluster, PTG_3h_perCluster, LDM_12h_perCluster, PTG_12h_perCluster, ctrl_perCluster, LDM_Lys3h_perCluster, PTG_Lys3h_perCluster, LPS_perCluster)
colnames(identpercluster) <- c("cluster", "LDM 3h", "PTG 3h", "LDM 12h", "PTG 12h", "control", "LDM Lysate", "PTG Lysate", "LPS")
identpercluster <- gather(identpercluster, key = "identity", value = "cellcount", -cluster)

#calculate percentages using an apply function 
percent <- function(x){
  ((x/sum(identpercluster$cellcount))*100)
}

identpercluster$percent <- identpercluster[,3]/(sum(identpercluster$cellcount))*100

#Plot results:
#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/originialIDpercluster_percent.png", width = 20, height = 10, unit = "cm", res = 300)
print(ggplot(identpercluster, aes(cluster, percent, fill= identity)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("") +
  ylab ("relative proportion of cells") + 
  scale_fill_manual(values = c("blue3", "dodgerblue", "olivedrab3", "red4", "magenta4", "deeppink3", "seagreen3", "forestgreen")) + 
  theme_classic() + 
  theme(legend.title = element_blank()))
#dev.off()

```


```{r, scale the calcualted percentages}
#calculate percentages of individual clusters with normalizing to the individual cluster 

c0.p <- subset(identpercluster, cluster == "cluster0")
c1.p <- subset(identpercluster, cluster == "cluster1")
c2.p <- subset(identpercluster, cluster == "cluster2")
c3.p <- subset(identpercluster, cluster == "cluster3")
c4.p <- subset(identpercluster, cluster == "cluster4")
c5.p <- subset(identpercluster, cluster == "cluster5")
c6.p <- subset(identpercluster, cluster == "cluster6")
c7.p <- subset(identpercluster, cluster == "cluster7")
c8.p <- subset(identpercluster, cluster == "cluster8")

cp.l <- c(c0.p$cellcount, c1.p$cellcount, c2.p$cellcount, c3.p$cellcount,c4.p$cellcount, c5.p$cellcount, c6.p$cellcount, c7.p$cellcount , c8.p$cellcount)


c0.p$percent.cluster <-  c0.p$cellcount/sum(c0.p$cellcount)*100
c1.p$percent.cluster <-  c1.p$cellcount/sum(c1.p$cellcount)*100
c2.p$percent.cluster <-  c2.p$cellcount/sum(c2.p$cellcount)*100
c3.p$percent.cluster <-  c3.p$cellcount/sum(c3.p$cellcount)*100
c4.p$percent.cluster <-  c4.p$cellcount/sum(c4.p$cellcount)*100
c5.p$percent.cluster <-  c5.p$cellcount/sum(c5.p$cellcount)*100
c6.p$percent.cluster <-  c6.p$cellcount/sum(c6.p$cellcount)*100
c7.p$percent.cluster <-  c7.p$cellcount/sum(c7.p$cellcount)*100
c8.p$percent.cluster <-  c8.p$cellcount/sum(c8.p$cellcount)*100

identpercluster1 <- dplyr::bind_rows(c0.p,c1.p, c2.p,c3.p, c4.p, c5.p, c6.p, c7.p, c8.p)

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/cellIDpercluster-norm.png", width = 20, height = 10, units = "cm", res = 300) 
  print(ggplot(identpercluster1, aes(cluster, percent.cluster, fill= identity)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("") +
  ylab ("cells of original condition per cluster [%]") + 
  scale_fill_manual(values = c("blue3", "dodgerblue", "olivedrab3", "red4", "magenta4", "deeppink3", "seagreen3", "forestgreen")) + 
  theme_classic() + 
  theme(legend.title = element_blank()))
#  dev.off()

```

Next we perform a variety of differential gene expression analysis based on multiple factors and conditions

1. Find markers based on the original cell class in the metadata, i.e. based on the experimental condition

```{r, Find markers for original identities}

se.dc.mus <- SetIdent(se.dc.mus, value = "cell_class")

dc.markers.original <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.5)

dc.markers.original <- subset(dc.markers.original, p_val_adj < 0.05)

#write.xlsx(dc.markers.original, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-orginal-IDs.xlsx")

#Visualize in a heatmap 

top20 <- dc.markers.original %>% group_by(cluster) %>% top_n(n=20, wt =avg_log2FC)
top.20 <- DoHeatmap(se.dc.mus, features = top20$gene, group.colors = c("dodgerblue", "blue3", "forestgreen", "olivedrab", "deeppink3", "magenta4", "seagreen3", "red4"))

top.20

```

2. Find markers for the identified clusters in the cluster analysis 

```{r, Find Markers according to cluster identity and subset them for later gprofiler analysis}

se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")

#Find markers for all clusters
se.dc.markers <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.5, only.pos = T)

#subset them for only significantly differential expressed genes 
se.dc.markers <- subset(se.dc.markers, p_val_adj < 0.05)

##visualize in a heatmap: 

top10 <- se.dc.markers %>% group_by(cluster) %>% top_n(n=10, wt =avg_log2FC)
top.10 <- DoHeatmap(se.dc.mus, features = top10$gene, group.colors = c.cols)
top.10

#look at differential expressed genes between the two clusters that comprise most of the 12h time-point (independent of strain, cluster 3 and cluster 5)
dc.12h.markers.3 <- FindMarkers(se.dc.mus, ident.1 = 3, ident.2 = 4, min.pct = 0.25, logfc.threshold = 0.5, only.pos = T)
dc.12h.markers.4 <- FindMarkers(se.dc.mus, ident.1 = 4, ident.2 = 3, min.pct = 0.25, logfc.threshold = 0.5, only.pos = T)

dc.12h.markers.3 <- subset(dc.12h.markers.3, p_val_adj < 0.05 & avg_log2FC > 2)
dc.12h.markers.4 <- subset(dc.12h.markers.4, p_val_adj < 0.05 & avg_log2FC > 2)
#visualize in a volcano plot
volcano12h.3.4 <- EnhancedVolcano(dc.12h.markers , lab = rownames(dc.12h.markers),x= "avg_log2FC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcano-12hDEG-leftPTG-rightLDM.png", width = 15, height = 25, units = "cm", res= 300)
print(volcano12h.3.4)
#dev.off()

#write.xlsx(dc.12h.markers,"~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/DEG-cluster3-cluster5.xlsx", row.names = T)

```

Classify the cells based on their expression profile using the CLustifyr package and the microarray data of [Helft J., 2015](https://pubmed.ncbi.nlm.nih.gov/26084029/). 



```{r, load library and format microarray reference data}

##use the Helft normalized expression from the limma analysis pipeline: data.matrix

#take the mean expression of each array and gene

data.mat1 <- data.frame(GM_DC = rowMeans(data.matrix[,1:3]), GM_Mac = rowMeans(data.matrix[,4:6]))

#translate the IDS to gene symbols

#merge ensembl name and gene ID
data.mat1 <- subset(data.mat1, rownames(data.mat1) %in% intersect(rownames(data.mat1), ann$ID))
data.mat1$ID <- rownames(data.mat1)
data.mat1 <- dplyr::left_join(data.mat1, ann, by = "ID")
#set merged new name as rownames for gene entries
dup <- which(duplicated(data.mat1$gene_symbol))
data.mat1 <- data.mat1[-dup,]
rownames(data.mat1) <- data.mat1$gene_symbol

data.mat <- data.mat1[,1:2]

```

```{r, prepare necessary data from the single cell data}

##workflow optimized for seurat objects
#run clustifyr on data

se.dc.mus <- clustifyr::clustify(input = se.dc.mus,ref_mat = data.mat, cluster_col = "seurat_clusters", obj_out = T)
#used number of genes: 774
#threshold = 0.5 

head(se.dc.mus[[]])


DimPlot(se.dc.mus, group.by = "celltype", cols = type.cols)

##make a column with cell_class and type in the object 

se.dc.mus$cell_class_type <- paste0(se.dc.mus$cell_class, "_", se.dc.mus$celltype)
se.dc.mus$cluster_type <- paste0(se.dc.mus$seurat_clusters, "_", se.dc.mus$celltype)
se.dc.mus$cluster_cell_class <- paste0(se.dc.mus$seurat_clusters, "_", se.dc.mus$cell_class)

DimPlot(se.dc.mus, group.by = "celltype", cols = type.cols)

#Visualize celltype signature in a heatmap

se.dc.mus <- SetIdent(se.dc.mus, value = "celltype" )
deg_celltype <- FindAllMarkers(se.dc.mus, logfc.threshold = 3)
deg_celltype <- subset(deg_celltype, deg_celltype$p_val_adj < 0.05)

cmat <- hmat_mousetoxo(t(se.dc.mus@assays$SCT@counts), se.dc.mus$cluster_type)


pdf("~/t.gondi_dc/res/figures/fig2/hmat_celltypes.pdf")
heatmap.F(cmat[deg_celltype$gene,], clustermethod = "ward.D2")
dev.off()

png("~/t.gondi_dc/res/figures/fig2/hmat_celltypes.png", height = 10, width = 15, res = 600)
heatmap.F(cmat[deg_celltype$gene,], clustermethod = "ward.D2")
dev.off()

deg_celltype <- deg_celltype[!duplicated(deg_celltype$gene), ]


pdf("~/t.gondi_dc/res/figures/fig2/hmat_celltypes.pdf")
phmat_simple(cmat, seed = unique(deg_celltype$gene))
dev.off()

```

Run cluster analysis separately on GM-DCs and GM-Macs 

```{r, separate raw data files for DCs and Macrophages and cluster them}

se.dc.mus <- SetIdent(se.dc.mus, value = "celltype")

gmDC.cells <- as.vector(colnames(subset(se.dc.mus, celltype == "GM-DC")))
gmMAC.cells <- as.vector(colnames(subset(se.dc.mus, celltype == "GM-Mac")))

gmDC.cnt <- se.dc.mus@assays$RNA@data[, gmDC.cells]

gmMAC.cnt <- se.dc.mus@assays$RNA@data[, gmMAC.cells]

#subset metadata

metaDC <- subset(meta1, rownames(meta1) %in% gmDC.cells)
metaMac <- subset(meta1, rownames(meta1) %in% gmMAC.cells)
  
se.dc <- CreateSeuratObject(gmDC.cnt, min.cells = 3, min.features = 200, project = "DC-TG-FH-M", meta.data = metaDC)
se.mac <- CreateSeuratObject(gmMAC.cnt, min.cells = 3, min.features = 200, project = "DC-TG-FH-M", meta.data = metaMac)


```

Now run the analysis for DCs only 

```{r, DC analysis}

se.dc <- SCTransform(se.dc)

se.dc <- RunPCA(se.dc)

print(se.dc[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(se.dc, dims = 1:2, reduction = "pca")

DimPlot(se.dc,  reduction = "pca", group.by = "cell_class", cols = "Dark2")

DimHeatmap(se.dc, dims = 1:15, cells = 500, balanced = TRUE)

ElbowPlot(se.dc, ndims= 20, reduction ="pca")

se.dc <- FindNeighbors(se.dc, dims = 1:6)
se.dc <- FindClusters(se.dc) #start with default resolution of 0.8
#inspect the clusters
head(Idents(se.dc), 5)

se.dc <- RunUMAP(se.dc, dims = 1:10, spread = 0.5, min.dist = 0.1)
#min.dist let's me play around with the thightnes of the cells within the clusters
#the spread let's me play around with the coordinates of the distance between the clusters 
##default settings are -> min.dist: 0.3, spread: 1

DimPlot(se.dc, reduction = "umap",group.by = "seurat_clusters")

DimPlot(se.dc, reduction = "umap", group.by = "cell_class")




```

```{r, Mac analysis}

se.mac <- SCTransform(se.mac)

se.mac <- RunPCA(se.mac)


print(se.mac[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(se.mac, dims = 1:2, reduction = "pca")

DimPlot(se.mac,  reduction = "pca", group.by = "cell_class", cols = "Dark2")

DimHeatmap(se.mac, dims = 1:15, cells = 500, balanced = TRUE)

ElbowPlot(se.mac, ndims= 20, reduction ="pca")

se.mac <- FindNeighbors(se.mac, dims = 1:15)
se.mac <- FindClusters(se.mac) #start with default resolution of 0.8
#inspect the clusters
head(Idents(se.mac), 5)

se.mac <- RunUMAP(se.mac, dims = 1:10, spread = 0.5, min.dist = 0.1)
#min.dist let's me play around with the thightnes of the cells within the clusters
#the spread let's me play around with the coordinates of the distance between the clusters 
##default settings are -> min.dist: 0.3, spread: 1

DimPlot(se.mac, reduction = "umap",group.by = "seurat_clusters")

DimPlot(se.mac, reduction = "umap", group.by = "cell_class")

#saveRDS(se.dc, "~/t.gondi_dc/res/seurat-objects/dc-only.RDS")

```


We also would like to run a cell cycle scoring analysis of the individual cell types to see whether Toxoplasma infected cells (particularly at the 12h time-point are at a specific changes in the cell cycle based on infection)

We got cell cycle associated gene lists based on he orthology analysis of Tirosh, I. et al ;  https://www-ncbi-nlm-nih-gov.ezp.sub.su.se/pmc/articles/PMC4944528/

```{r, cell cycle scoring - prepare data}

cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Mus_musculus.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

# Acquire the S phase genes
s_ids <- cell_cycle_genes %>%
        dplyr::filter(phase == "S") %>%
        pull("geneID")
        
# Acquire the G2M phase genes        
g2m_ids <- cell_cycle_genes %>%
        dplyr::filter(phase == "G2/M") %>%
        pull("geneID")


an <- genes.table[cell_cycle_genes$geneID,]

an <- data.frame(geneID = genes.table$ensembl_gene_id, gene_name = genes.table$external_gene_name)

cell_cycle_markers <- dplyr::left_join(cell_cycle_genes, an, by = "geneID")

s_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "S") %>%
        pull("gene_name")
        
# Acquire the G2M phase genes        
g2m_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "G2/M") %>%
        pull("gene_name")


```

```{r, cell cycle scoring - run scoring}

se.dc <- CellCycleScoring(se.dc, g2m.features = g2m_genes, s.features = s_genes)
se.mac <- CellCycleScoring(se.mac, g2m.features = g2m_genes, s.features = s_genes)


DimPlot(se.dc, reduction = "pca", group.by = "Phase")
DimPlot(se.dc, reduction = "umap", group.by = "Phase")

DimPlot(se.mac, reduction = "pca", group.by = "Phase")
DimPlot(se.mac, reduction = "umap", group.by = "Phase")

se.dc.mus <- CellCycleScoring(se.dc.mus, g2m.features = g2m_genes, s.features = s_genes)
DimPlot(se.dc.mus, reduction = "umap", group.by = "Phase")

#png("~/t.gondi_dc/res/figures/fig2/cell_cycle_scoring.png",width = 17.5, height = 15, res = 600, units = "cm")
DimPlot(se.dc.mus, reduction = "umap", group.by = "Phase", cols = m.cycle.c) +
                              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank())
#dev.off()

```
This observation here is quite interesting : The cell cycle does not seem like we have to regress it out as in the PCA we can see a trend that G1 and S-Phase are most responsible for a relatively high degree of variation, as we are interested in this difference especially in light of the fact that S-phase annotated genes are mostly cells with a 12h infection we don't want to regress this co-variate but rather investigate it further. 


Based on doi:10.1111/j.1462-5822.2008.01117.x , T. gondii (RH) infection promotes entry into the S phase

```{r, visualization of cell cycle regulators across clusters, fig.width= 20, fig.height= 4}


DotPlot(se.dc.mus, dot.scale = 25, cols = c("#440154", "#FDE725"),features =  c("Ccna1", "Ccnb1","Ccnd1", "Ccnd2", "Ccnd3", "Ccne1", "Ccne2","Cdk2", "Cdk1", "Cdk4", "Cdk6"))

# Create a data.frame with one column for continuous values, e.g. gene expression
# and one column with a group variable, e.g. clusters

cyclins <- c("Ccna1", "Ccnb1","Ccnd1", "Ccnd2", "Ccnd3", "Ccne1", "Ccne2")
cdks <- c("Cdk2", "Cdk1", "Cdk4", "Cdk6")
cdkInh <- c("Cdkn1a","Cdkn2c", "Cdkn1b","Cdkn1c","Cdkn2a","Cdkn2b")
cdkreg <- c("Cdc25a", "Cks1b","Cks2", "Gadd45a")

#cluster = vector with cell and clusterannotation/celltype annotation or annotation of interest
#gene = vector with expression values for a gene of interest in the data
#dataset = which dataset to use (dc, dc+mac,mac)
#cols = colors used for the clusters 

CyCDK_expr <- function(object, clust, gene, cols){
  #cyclin B
gg <- data.frame(cluster = clust, val = object@assays$SCT@data[gene,])
# x should be the continuous variable and y the categorical group variable
# You can color the ridges using any variable you want
  gg <- ggplot(gg, aes(x = val, y = cluster, fill = cluster)) +
  geom_density_ridges(alpha = 0.8) +# Set transparency
  scale_fill_manual(values = cols) +
  theme_ridges() +
  labs(x = "Expression value", y = "Cluster identity", fill = "Cluster identity") + 
  ggtitle(gene)
  return(gg)
}

#cyclin B
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$seurat_clusters, gene = c("Ccnb1"), cols = c.cols)
#cyclin D1
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$seurat_clusters, gene = "Ccnd1",cols = c.cols)
#cyclin D2
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Ccnd2", cols = cell.cols)
#cyclin D3
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Ccnd3", cols = cell.cols)
#cyclin E1
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Ccne1", cols = cell.cols)
#cyclin E2
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$seurat_clusters, gene = "Ccne2", cols = c.cols)


png("~/t.gondi_dc/res/figures/fig2/cyclin_1_to_3_density.png", res = 600, width = 50, height = 10, units = "cm")
cowplot::plot_grid(cycp1, cycp2, cycp3, cols = 3)
dev.off()

png("~/t.gondi_dc/res/figures/fig2/cyclin_4_to_6_density.png", res = 600, width = 50, height = 10, units = "cm")
cowplot::plot_grid(cycp4, cycp5, cycp6, cols = 3)
dev.off()

cdks <- c("Cdk2", "Cdk1", "Cdk4", "Cdk6")
#cdk1
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Cdk1", cols = cell.cols)
#cdk2
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Cdk2", cols = cell.cols)
#cdk4
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Cdk4", cols = cell.cols)
#cdk6
CyCDK_expr(object = se.dc.mus, clust = se.dc.mus$cell_class, gene = "Cdk6", cols = cell.cols)


png("~/t.gondi_dc/res/figures/fig2/cdk_density.png", res = 600, width = 50, height = 10, units = "cm")
cowplot::plot_grid(cdk1, cdk2, cdk3, cdk6,cols = 4)
dev.off()

```


```{r, run GO term GO term analyis based on clusters}

dc.markers.C0 <- subset(se.dc.markers, se.dc.markers$cluster == "M 1")
dc.markers.C1 <- subset(se.dc.markers, se.dc.markers$cluster == "M 2")
dc.markers.C2 <- subset(se.dc.markers, se.dc.markers$cluster == "M 3")
dc.markers.C3 <- subset(se.dc.markers, se.dc.markers$cluster == "M 4")
dc.markers.C4 <- subset(se.dc.markers, se.dc.markers$cluster == "M 5")
dc.markers.C5 <- subset(se.dc.markers, se.dc.markers$cluster == "M 6")
dc.markers.C6 <- subset(se.dc.markers, se.dc.markers$cluster == "M 7")
dc.markers.C7 <- subset(se.dc.markers, se.dc.markers$cluster == "M 8")
dc.markers.C8 <- subset(se.dc.markers, se.dc.markers$cluster == "M 9")


marker.list <- list(dc.markers.C0, dc.markers.C1, dc.markers.C2, dc.markers.C3, dc.markers.C4, dc.markers.C5, dc.markers.C6, dc.markers.C7, dc.markers.C8)

dc.markers.n <- c("dc.markers.C0","dc.markers.C1", "dc.markers.C2", "dc.markers.C3","dc.markers.C4", "dc.markers.C5", "dc.markers.C6", "dc.markers.C7", "dc.markers.C8")

subset.markers.l <- setNames(lapply(marker.list, function(x) {
  subset(x, avg_log2FC > 0.5 & p_val_adj < 0.01)
}), nm = dc.markers.n)

```


```{r, run GO term analysis with different databases}

library(gprofiler2)

gostres.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x[,7], organism = "mmusculus")
}), nm = c("gostres.C0", "gostres.C1", "gostres.C2", "gostres.C3", "gostres.C4", "gostres.C5", "gostres.C6","gostres.C7","gostres.C8"))

#subset for GO:BP only
gostres.BP.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x[,7], organism = "mmusculus", sources = "GO:BP")
}), nm = c("gostres.BP.C0", "gostres.BP.C1", "gostres.BP.C2", "gostres.BP.C3", "gostres.BP.C4", "gostres.BP.C5", "gostres.BP.C6", "gostres.BP.C7", "gostres.BP.C8"))

#subset for Kegg only

gostres.KEGG.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x[,7], organism = "mmusculus", sources = "KEGG")
}), nm = c("gostres.KEGG.C0", "gostres.KEGG.C1", "gostres.KEGG.C2", "gostres.KEGG.C3", "gostres.KEGG.C4", "gostres.KEGG.C5", "gostres.KEGG.C6", "gostres.KEGG.C7"))

#subset for reactome only

gostres.REAC.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x[,7], organism = "mmusculus", sources = "REAC")
}), nm = c("gostres.BP.C0", "gostres.BP.C1", "gostres.BP.C2", "gostres.BP.C3", "gostres.BP.C4", "gostres.BP.C5", "gostres.BP.C6", "gostres.BP.C7", "gostres.BP.C8"))



png("~/Desktop/Reac-C8.png", width = 30, height = 15, units = "cm", res = 300)
print(ggplot(head(gostres.REAC.l$gostres.BP.C8$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "darkblue") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 20, name = "Reds")) + 
  scale_y_continuous() +
  xlab("Biological Process")+
  ylab("Enrichment Score") +
  theme(axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.background = element_blank()))
dev.off()


GO.all.1 <- cowplot::plot_grid(p.go.0,NULL,p.go.4,NULL,p.go.6, ncol= 5, nrow = 1, rel_widths = c(5,0.1,5,0.1,5))
GO.all.1

GO.all.2 <- cowplot::plot_grid(p.go.1,NULL,p.go.2, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.2
GO.all.3 <- cowplot::plot_grid(p.go.3,NULL,p.go.5, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.3
GO.all.4 <- cowplot::plot_grid(p.go.7,NULL,p.go.8, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.4

#Save the plots all together 
#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/GO-enrichment/GO-enrichments-clusterbased-4.png", width = 60, height = 10, res = 300, units = "cm") #Units are pixels by default
print(GO.all.4)
#dev.off()

#Save the plots individually

#png(filename = "~/Desktop/PhDprojectFranziskaHildebrandt/TG-DC-Analysis/Results-200507-FiguresAndTables/GO-enrichment/GO-enrichment-C0.png", width = 30, height = 15, res = 300, units = "cm") #Units are pixels by default
print(p0)
#dev.off()

```

In addition: Try to run the DGEA according to the original cell-identity and look at the results and the GO-term analysis in the DCs! 


```{r, Clusterprofiler analysis for cluster 4}


markers_4 <- FindMarkers(se.dc.mus, ident.1 = "M 7") #refers to LDM infection at 12h
markers_4 <- subset(markers_4, p_val_adj < 0.05)
original.gene.list <- markers_4$avg_log2FC
markers_4$gene <- rownames(markers_4)
names(original.gene.list) <- markers_4$gene
gene_list<-na.omit(original.gene.list)
gene_list = sort(gene_list, decreasing = TRUE)
head(gene_list)

gseM7 <- gseGO(geneList=gene_list, 
             ont ="BP",
             keyType = "SYMBOL",
             minGSSize = 3, 
             maxGSSize = 500, 
             pvalueCutoff = 0.1, 
             pAdjustMethod = "hochberg",
             verbose = TRUE, 
             OrgDb = org.Mm.eg.db)

head(gseM7@result)


#select the most up/downregulated pathways/pathways of interest

ridgeplot(gse4, showCategory = 20,  core_enrichment = T) + 
  ggplot2::labs(x = "enrichment distribution") +
  ggplot2::scale_fill_gradientn(colors = colorRampPalette(c("#440154" ,"#21908C", "#FDE725"))(200))

#png("~/t.gondi_dc/res/figures/fig3/gse_genes_infected12h_enrichment.png", width = 30, height = 25, res = 600, units = "cm")
dotplot(gse4, showCategory= 10, split=".sign") + 
  ggplot2::facet_grid(.~.sign) +
  ggplot2::scale_color_gradientn(colors = colorRampPalette(c("#FDE725", "#21908C","#440154" ))(200))
#dev.off()


cnetplot(gse4, categorySize="pvalue",foldChange = gene_list)

p <- heatplot(gseM7, showCategory = 15, foldChange = gene_list) +
  ggplot2::scale_fill_gradientn(colors = colorRampPalette(c("#440154" ,"#21908C", "#FDE725"))(200))

edox2 <- pairwise_termsim(gseM6)
treeplot(edox2) + 
  scale_color_gradientn("p-value",colors = colorRampPalette(c("#660156" ,"#21908C", "#FDE725"))(200))

edox <- pairwise_termsim(gseM7)
treeplot(edox) + 
  scale_color_gradientn("p-value",colors = colorRampPalette(c("#660156" ,"white", "#FDE725"))(200))


#png("~/t.gondi_dc/res/figures/fig3/gse_genes_12h.png", width = 25, height = 12, units = "cm", res = 600)
print(p)
#dev.off()


```


```{r, Clusterprofiler analysis for cluster M2 (late infected dendritic cells)}

se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")
markers_m2 <- FindMarkers(se.dc.mus, ident.1 = "M 2", logfc.threshold = 0.25)
markers_m2 <- subset(markers_m2, p_val_adj < 0.05)
original.gene.list <- markers_m2$avg_log2FC
markers_m2$gene <- rownames(markers_m2)
names(original.gene.list) <- markers_m2$gene
gene_list<-na.omit(original.gene.list)
gene_list = sort(gene_list, decreasing = TRUE)
head(gene_list)

gseM2 <- gseGO(geneList=gene_list, 
             ont ="BP", 
             keyType = "SYMBOL",
             exponent = 1, 
             minGSSize = 3,
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = org.Mm.eg.db, 
             pAdjustMethod = "hochberg", 
             by = "fgsea")


#supplement add treeplot with all GO-terms 
edox2 <- pairwise_termsim(gseM2)

#png("~/t.gondi_dc/res/figures/fig3/treeplot_M2.png", width = 47, height = 25, units = "cm", res = 600)
treeplot(edox2) + 
  scale_color_gradientn("p-value",colors = colorRampPalette(c("#660156" ,"#21908C", "#FDE725"))(200))
#dev.off()


#Manual selection of terms of interest
gseM2@result <- subset(gseM2@result, gseM2@result$Description %in% c("DNA metabolic process",
                                                                     "regulation of cell motility",
                                                                     "defense response to bacterium", 
                                                                     "positive regulation of cytokine production", 
                                                                     "cell communication"
  
))



pdf("~/t.gondi_dc/res/figures/fig3/gse_genes_cluster6_enrichment.pdf")
ridgeplot(gseM2) + 
  ggplot2::labs(x = "enrichment distribution") +
  ggplot2::scale_fill_gradientn(colors = colorRampPalette(c("#FDE725", "#21908C" ,"#660156" ))(200)) + 
  ggplot2::scale_x_continuous(breaks=seq(-6, 6, 2)) +
  theme_classic() + 
  theme(axis.text.y = element_text( colour="black",size=10), 
        axis.text.x = element_text( colour="black",size=10))
dev.off()

#supplement
pdf("~/t.gondi_dc/res/figures/fig3/gse_genes_cluster_M2.pdf")
heatplot(gseM2, showCategory = 20, foldChange = gene_list) +
  ggplot2::scale_fill_gradientn(colors = colorRampPalette(c("#660156" ,"#21908C", "#FDE725"))(200))+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.text.y = element_text(size = 12))
dev.off()



```

```{r, look specifically into DEG between cluster 3 (12h_PTG) and cluster 4 (12h_LDM)}

#look at differentially expressed genes between the different infection clusters 
#M7 = LDM, M6 = PTG (both at 12h)
PTG_LDM_markers <- FindMarkers(se.dc.mus, ident.1 = "M 6", ident.2 = "M 7", logfc.threshold = 0.25)
PTG_LDM_markers <- subset(PTG_LDM_markers, PTG_LDM_markers$p_val_adj < 0.05)
PTG_LDM_markers <- PTG_LDM_markers[order(-PTG_LDM_markers$avg_log2FC),]
PTG_LDM_markers$gene <- rownames(PTG_LDM_markers)

PTG_markers <- subset(PTG_LDM_markers,PTG_LDM_markers$avg_log2FC > 0)
LDM_markers <- subset(PTG_LDM_markers,PTG_LDM_markers$avg_log2FC < 0)

#write.table(rownames(PTG_markers), "~/t.gondi_dc/res/DEG/PTG_markers_180121.txt", quote = F, col.names = F, row.names = F)

#heatmap
p_all <- hmat_mousetoxo(data = t(se.dc.mus@assays$SCT@counts),
               meta = se.dc.mus$seurat_clusters)

pdf("~/t.gondi_dc/res/figures/fig3/heatmap_M6_M7.pdf")
phmat_simple(hmat = p_all, seed = PTG_LDM_markers$gene,
               cond = c(6,7))
dev.off()

heatmap.F(p_all[,c(6,7)], colors= colorRampPalette(c("#660156","#21908C", "white", "#FFF68F", "#FFD700"))(1001))

#run gsea
original.gene.list <- PTG_LDM_markers$avg_log2FC
PTG_LDM_markers$gene <- rownames(PTG_LDM_markers)
names(original.gene.list) <- PTG_LDM_markers$gene
gene_list<-na.omit(original.gene.list)
gene_list = sort(gene_list, decreasing = TRUE)
head(gene_list)

PTG_LDM_gsea <- gseGO(geneList=gene_list, 
             ont ="BP", 
             keyType = "SYMBOL",
             exponent = 1, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = org.Mm.eg.db, 
             pAdjustMethod = "none")

head(PTG_LDM_gsea@result)

edox2 <- pairwise_termsim(PTG_LDM_gsea)
#suppplement

#png("~/t.gondi_dc/res/figures/fig3/treeplot_M6_M7.png", width = 47, height = 25, units = "cm", res = 600)
treeplot(edox2)  + 
  scale_color_gradientn("p-value",colors = colorRampPalette(c("#660156","#21908C", "#FFF68F", "#FFD700"))(201), 
                        )
#dev.off()

PTG_LDM_gsea@result <- subset(PTG_LDM_gsea@result, PTG_LDM_gsea@result$Description %in% c("negative regulation of leukocyte activation","response to interferon-beta","response to bacterium","positive regulation of hydrolase activity","negative regulation of cell motility","positive regulation of cysteine-type endopeptidase activity"))


pdf("~/t.gondi_dc/res/figures/fig3/heatplot_M6_M7.pdf")
heatplot(PTG_LDM_gsea, foldChange = gene_list) +
  ggplot2::scale_fill_gradientn(colors = colorRampPalette(c("#660156","#21908C", "white", "#FFF68F", "#FFD700"))(201)) +
  theme_classic() + 
  theme(axis.text.y = element_text( colour="black",size=12), 
        axis.text.x = element_text( colour="black",size=10, angle = 45, hjust = 1)) +
  coord_flip()
dev.off()

#get the core enrichment genes as a vector to double check where they are in the heatmap (plot only these)
core_PTG_LDM_gsea <- PTG_LDM_gsea@result$core_enrichment
core_PTG_LDM_gsea <- unique(unlist(str_split(core_PTG_LDM_gsea, "/")))


print(p_all)

p_core <- hmat_mousetoxo(data = mat,
               meta = se.dc.mus$seurat_clusters, 
               seed = core_PTG_LDM_gsea,
               cond = c(4,7))

print(p_core)

```

Heatplot shows genes in alphabetical order but I would like to have them ordered by either association with the go-term or customized, thus I will write my own function to plot these accordingly: 

```{r, custom heatplot}

#extract GO-terms and enrichment in new dataframe

go.table <- subset(PTG_LDM_gsea@result, PTG_LDM_gsea@result$Description %in% c("negative regulation of leukocyte activation","response to interferon-beta","response to bacterium","positive regulation of hydrolase activity","negative regulation of cell motility","positive regulation of cysteine-type endopeptidase activity"))

go.table

#make data frame with 1 column as the description and one column with all the genes

nrow(go.table)

l <- lapply(1:nrow(go.table), function(x){
  df <- data.frame(genes = do.call(cbind, strsplit(go.table[x,11], "/", fixed = T)), goterm= paste(go.table[x,2]))
  })

library(data.table)

df <- rbindlist(l)

#now we can add the logFCs from our genelist

logFC <- as.data.frame(subset(PTG_LDM_gsea@geneList, names(PTG_LDM_gsea@geneList) %in%  df$genes))
logFC$genes <- rownames(logFC)
colnames(logFC) <- c("logFC", "genes")
head(logFC)

#now combine the data 

library(dplyr)

df <- left_join(df, logFC, by = "genes")
df$genes <- as.factor(df$genes)

p <- df %>%
  mutate(genes = fct_relevel(genes, rev(c(
            "Ifi203", "Naip2", "Il1b", 
            "Axl", "Abca1", "Cd14", 
            "Ikbke", "Fpr2","Clec4d","Mndal", "Cxcl3", "Clec4e", "Acod1","Socs1","Lag3", "Ripk2", "Pparg", "Ccl24", "Cdh1","Hfe", "Cx3cl1", "Fyn", "Hip1"))))

# now we can plot the data in ggplot

pdf("~/t.gondi_dc/res/figures/fig3/heatplot_M6_M7.pdf")
ggplot(p, aes(genes, goterm)) +
  geom_tile(aes(fill = logFC), colour = "white") + 
  scale_fill_gradientn(breaks = seq(round(-max(abs(p$logFC))), round(max(abs(p$logFC))), by = 0.5), limits = c(min(p$logFC),max(p$logFC)), colors = colorRampPalette(c("#0000E5","#9999FF", "white", "#F2A48C", "#ff4d04"))(41)) + 
  coord_flip()+
  xlab("gene")+ 
  ylab("enriched term (BP)") +
  theme(panel.background = element_rect(fill = "white"), axis.text.x = element_text(angle = 35, hjust = 1), axis.line = element_line(size=0.2, color = "black"))
dev.off()
```


There is a distinction between mature and immature DCs [Shalek A., 2013](https://pubmed.ncbi.nlm.nih.gov/23685454/) or GM-DCs (= mature) and GM-Macs (= immature) [Helft J., 2015](https://pubmed.ncbi.nlm.nih.gov/26084029/). Therefore we want to see whether the identified sub-populations in our UMAP correspond to these classifications by using markers that correspond to these classifications.

```{r, compare gene expression in all clusters, fig.width= 18, fig.height=5}
#First convert upregulated genes in a list for each cluster
clusters <- c("M 1", "M 2", "M 3", "M 4", "M 5", "M 6", "M 7", "M 8", "M 9")

go_list <- setNames(lapply(clusters, function(x){
  subset(se.dc.markers[,7],se.dc.markers[,6] == x)
  }), nm = clusters)

cell_class <- c( "ctrl","LDM_12h","LDM_3h","LDM_Lys3h","LPS", "PTG_12h","PTG_3h","PTG_Lys3h")
#do the same for original conditions 
go_list_2 <- setNames(lapply(cell_class, function(x){
  subset(dc.markers.original[,7],dc.markers.original[,6] == x)
  }), nm = cell_class)

ck <- compareCluster(geneClusters = go_list,
                     fun = "enrichGO",
                     keyType = "SYMBOL", 
                     OrgDb = org.Mm.eg.db,
                     pvalueCutoff = 0.01,
                     pAdjustMethod = "hochberg")

png("~/t.gondi_dc/res/figures/fig2/GO_clusters_compare.png", width = 55, height = 12, res = 600, unit = "cm")
dotplot(ck) + 
  scale_color_gradientn("adj.p-value",colors = colorRampPalette(c("#FDE725", "#21908C","#660156"))(200)) +
  theme(panel.background = element_rect(fill = c("#7F7F7F")), panel.grid = element_blank(), axis.text.y = element_text(size = 10), axis.text.x = element_text(size = 10, angle = 60, hjust = 1))+ 
  coord_flip()
dev.off()
  

```
Integrate toxodata in the mouse data 


```{r, add toxoclusters to metadata and highlight them in umap}

toxo_clusters <- data.frame(toxo.cluster = paste0("T", " ", se.toxo$seurat_clusters), cell = rownames(se.toxo[[]]))
rownames(toxo_clusters) <- toxo_clusters$cell
names(toxo_clusters)[1] <- "toxo_cluster"
toxo_clusters$cell <- NULL
head(toxo_clusters)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_clusters, col.name = "toxoclusters")

#Visulize only cells of toxo_cluster = toxo_2 on DC uMAP!

DimPlot(se.dc.mus, group.by = "toxoclusters", split.by = "toxoclusters")

toxo.dc.0 <- DimPlot(se.dc.mus, group.by = "toxoclusters", cols = c("darkred", "gray48", "gray48", "gray48", "gray48", "gray48", "gray48"))

toxo.dc.1 <- DimPlot(se.dc.mus, group.by = "toxoclusters", cols = c("gray48", "darkred", "gray48", "gray48", "gray48", "gray48", "gray48"))

toxo.dc.2 <- DimPlot(se.dc.mus, group.by = "toxoclusters", cols = c("gray48", "gray48","darkred" ,"gray48", "gray48", "gray48", "gray48"))
  
toxo.dc.3 <- DimPlot(se.dc.mus, group.by = "toxoclusters", cols = c("gray48",  "gray48", "gray48", "darkred","gray48", "gray48", "gray48"))


#Supplement
#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/Toxo-Clusters-DC-Umap.png", width = 40, height = 25, unit= "cm", res= 300)
print(cowplot::plot_grid(toxo.dc.0, toxo.dc.1, toxo.dc.2, toxo.dc.3))
#dev.off()

D
  
```

```{r, add toxo cell cycle as metadata to the mouse data}

toxo_phase <- as.data.frame(paste0(se.toxo$Phase))
rownames(toxo_phase) <- rownames(se.toxo[[]])
names(toxo_phase)[1] <- "toxo_phase"
head(toxo_phase)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_phase, col.name = "toxo_phase")

DimPlot(se.dc.mus, group.by = "toxo_phase")

```


```{r, visualize the cells that are Gra15+ and Rop16+ in the DC set}

toxo_gra <- as.data.frame(paste0(se.toxo$Gra15_expression))
rownames(toxo_gra) <- rownames(se.toxo[[]])
names(toxo_gra)[1] <- "gra-expression"
head(toxo_gra)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_gra, col.name = "Gra15_expression")


toxo_rop <- as.data.frame(paste0(se.toxo$Rop16_expression))
rownames(toxo_rop) <- rownames(se.toxo[[]])
names(toxo_rop)[1] <- "rop-expression"
head(toxo_rop)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_rop, col.name = "Rop16_expression")

DimPlot(se.dc.mus, group.by = "Rop16_expression")

se.dc.rop <- subset(se.dc.mus, subset = Rop16_expression %in% "Rop16+", slot = "data")
se.dc.gra <- subset(se.dc.mus, subset = Gra15_expression %in% "Gra15+", slot = "data")
DimPlot(se.dc.rop, group.by = "toxoclusters")
DimPlot(se.dc.gra, group.by = "cell_class")

DimPlot(se.dc.mus, group.by = "Gra15_expression")
DimPlot(se.dc.mus, group.by = "Rop16_expression")

```

```{r, heatmap between cell-types, visualization of signature}

Idents(se.dc.mus) <- "cell_class_type"
lps <- WhichCells(se.dc.mus, idents = c("LPS_GM_Mac","LPS_GM_DC")) #subset cluster 2 
lps_ctrlset <- subset(se.dc.mus, idents = c("ctrl_GM_Mac", "ctrl_GM_DC"))
dc_set <- subset(se.dc.mus, idents = c("LPS_GM_DC","ctrl_GM_DC", "PTG_12h_GM_DC", "PTG_3h_GM_DC", "LDM_3h_GM_DC", "LDM_12h_GM_DC","PTG_Lys3h_GM_DC", "LDM_Lys3h_GM_DC"))
mac_set <- subset(se.dc.mus, idents = c("LPS_GM_Mac", "ctrl_GM_Mac", "PTG_12h_GM_Mac", "PTG_3h_GM_Mac", "PTG_Lys3h_GM_Mac", "LDM_3h_GM_Mac", "LDM_12h_GM_Mac","LDM_Lys3h_GM_Mac"))
```


```{r, make new metadata}


meta_mcp <- data.frame(pathology = as.factor(ifelse(se.dc.mus@meta.data$cell_class %in% c("LDM_3h","LDM_12h","PTG_3h","PTG_12h"), paste("infected"), ifelse(se.dc.mus@meta.data$cell_class %in% c("LPS"), paste("LPS"), ifelse(se.dc.mus@meta.data$cell_class %in% c("PTG_Lys3h", "LDM_Lys3h"), paste("Lysate"),paste("uninfected"))))),
                       cell_ID = rownames(se.dc.mus@meta.data),
                       timepoint = as.factor(ifelse(se.dc.mus@meta.data$cell_class %in% c("control","LDM_Lys3h","PTG_Lys3h", "LPS"),paste(0),ifelse(se.dc.mus@meta.data$cell_class %in% c("LDM_3h", "PTG_3h"), paste(3),paste(12)))), 
                       p_toxo = as.numeric(se.dc.mus@meta.data$toxo_nUMI_p), 
                       phase = as.factor(se.dc.mus@meta.data$Phase),
                       sample = as.factor(se.dc.mus@meta.data$cell_class), 
                       strain = as.factor(ifelse(se.dc.mus@meta.data$cell_class %in% c("LDM_3h", "LDM_Lys3h", "LDM_12h"), paste("LDM"), ifelse(se.dc.mus@meta.data$cell_class %in% c("PTG_3h", "PTG_Lys3h", "PTG_12h"), paste("PTG"), paste("None")))),
                       cellQ = as.numeric(log10(se.dc.mus@meta.data$nFeature_SCT)))

```

We want to run RNA velocity experiments using scvelo in python - to this end we are exporting the metadata of our seurat object and format it into an adata object using this tutorial: https://smorabit.github.io/tutorials/8_velocyto/

```{r, write an adata object for velocity experiments}

se.dc.mus$barcode <- colnames(se.dc.mus)
se.dc.mus$UMAP_1 <- se.dc.mus@reductions$umap@cell.embeddings[,1]
se.dc.mus$UMAP_2 <- se.dc.mus@reductions$umap@cell.embeddings[,2]

# save metadata table:
write.csv(se.dc.mus@meta.data, file = "~/t.gondi_dc/data/adata_create/metadata.csv", quote = F, row.names = F)

# save metadata table:library(Matrix)
counts_matrix <- GetAssayData(se.dc.mus, assay = "RNA", slut = "counts")

# write expression counts matrix
writeMM(counts_matrix, file= "~/t.gondi_dc/data/adata_create/counts.mtx")

# write dimensionality reduction matrix, in this example case pca matrix
write.csv(se.dc.mus@reductions$pca@cell.embeddings, file="~/t.gondi_dc/data/adata_create/pca.csv",quote=F, row.names=F)

# write gene names
write.table(
  data.frame('gene'=rownames(counts_matrix)),file='~/t.gondi_dc/data/adata_create/gene_names.csv',
  quote=F,row.names=F,col.names=F
)

```

We would like to compare the immune response of Toxoplasma infected DCs to the response induced by other well-established ligands for pathogen-sensing pathways. 

1. LPS (lipopolysaccharide) > TLR4
2. PAM (Pam3CSK4) > TLR2
3. high molecular weight poly(I:C) (HMW) > TLR3/MDA-5
4. CpG DNA type B (cpGB) > TLR9 
5. Sendai virus (SeV) > RIG-I
6. depleted zymosan (zymosan) > Dectin-1
7. cyclic [G(3'-5')pA(3'-5')p] (cGAMP) > STING

```{r, make gene lists}

gene.set_LPS <- read.table("~/t.gondi_dc/res/GEO_comparison/LPS_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_LPS <- gene.set_LPS$x
gene.set_LPS <- subset(gene.set_LPS, gene.set_LPS %in% rownames(se.dc.mus))

gene.set_PAM <- read.table("~/t.gondi_dc/res/GEO_comparison/PAM_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_PAM <- gene.set_PAM$x
gene.set_PAM <- subset(gene.set_PAM, gene.set_PAM %in% rownames(se.dc.mus@assays$RNA))

gene.set_HMW <- read.table("~/t.gondi_dc/res/GEO_comparison/HMW_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_HMW <- gene.set_HMW$x
gene.set_HMW <- subset(gene.set_HMW, gene.set_HMW %in% rownames(se.dc.mus@assays$RNA))

gene.set_Sev <- read.table("~/t.gondi_dc/res/GEO_comparison/SeV_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_Sev <- gene.set_Sev$x
gene.set_Sev <- subset(gene.set_Sev, gene.set_Sev %in% rownames(se.dc.mus@assays$RNA))

gene.set_cGAMP <- read.table("~/t.gondi_dc/res/GEO_comparison/cGAMP_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_cGAMP <- gene.set_cGAMP$x
gene.set_cGAMP <- subset(gene.set_cGAMP, gene.set_cGAMP %in% rownames(se.dc.mus@assays$RNA))

gene.set_cpGB <- read.table("~/t.gondi_dc/res/GEO_comparison/cpGB_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_cpGB <- gene.set_cpGB$x
gene.set_cpGB <- subset(gene.set_cpGB, gene.set_cpGB %in% rownames(se.dc.mus@assays$RNA))

gene.set_zymosan <- read.table("~/t.gondi_dc/res/GEO_comparison/Zymosan_pos_log2FC.tsv", sep = "\t", header = T)
gene.set_zymosan <- gene.set_zymosan$x
gene.set_zymosan <- subset(gene.set_zymosan, gene.set_zymosan %in% rownames(se.dc.mus@assays$RNA))


```

```{r, extract unique genes for each condition, including infections}

Sev_unique <- setdiff(gene.set_Sev, c(gene.set_cGAMP,gene.set_cpGB, gene.set_LPS, gene.set_HMW, gene.set_zymosan, gene.set_PAM))
zymosan_unique <- setdiff(gene.set_zymosan, c(gene.set_cGAMP,gene.set_cpGB, gene.set_LPS, gene.set_HMW, gene.set_Sev, gene.set_PAM))
cGAMP_unique <- setdiff(gene.set_cGAMP, c(gene.set_zymosan,gene.set_cpGB, gene.set_LPS, gene.set_HMW, gene.set_Sev, gene.set_PAM))
LPS_unique <- setdiff(gene.set_LPS, c(gene.set_zymosan,gene.set_cpGB, gene.set_cGAMP, gene.set_HMW, gene.set_Sev, gene.set_PAM))
HMW_unique <- setdiff(gene.set_HMW, c(gene.set_zymosan,gene.set_cpGB, gene.set_cGAMP, LPS_unique, gene.set_Sev, gene.set_PAM))
Pam_unique <- setdiff(gene.set_PAM, c(gene.set_zymosan,gene.set_cpGB, gene.set_cGAMP, LPS_unique, gene.set_Sev, gene.set_HMW))
cpGB_unique <- setdiff(gene.set_cpGB, c(gene.set_zymosan,gene.set_PAM, gene.set_cGAMP, LPS_unique, gene.set_Sev, gene.set_PAM))

```

```{r, add Average expression of gene sets of interest to the data }

LPS_rows <- which(rownames(se.dc.mus@assays$RNA) %in% gene.set_LPS)
# Get mean expression of genes of interest per cell
mean.exp <- colMeans(x = se.dc.mus@assays$RNA[LPS_rows,], na.rm = TRUE)
# Add mean expression values in 'object@meta.data$gene.set.score'
if (all(names(x = mean.exp) == rownames(x = se.dc.mus@meta.data))) {
  cat("Cell names order match in 'mean.exp' and 'object@meta.data':\n", 
      "adding gene set mean expression values in 'object@meta.data$gene.set.score'")
 se.dc.mus@meta.data$LPS_score <- mean.exp
}
# plot
png("~/t.gondi_dc/res/GEO_comparison/LPS_score_UMAP.png", width = 15, height = 10, units = "cm", res = 600)
FeaturePlot(object = se.dc.mus, 
            features = "LPS_score", pt.size = 2) +
                              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank()) + 
scale_color_gradientn("mean expression", colors = c("darkblue", "mediumblue", "yellow3", "yellow2", "yellow1", "yellow"))
dev.off()

```

```{r, vsualize gene of interests in UMAP (not included in manuscript)}
###Do the same for PAM specific expression 

PAM_rows <- which(rownames(se.dc.mus@assays$RNA) %in% gene.set_PAM)

# Get mean expression of genes of interest per cell
mean.exp <- colMeans(x = se.dc.mus@assays$RNA[PAM_rows,], na.rm = TRUE)

# Add mean expression values in 'object@meta.data$gene.set.score'
if (all(names(x = mean.exp) == rownames(x = se.dc.mus@meta.data))) {
  cat("Cell names order match in 'mean.exp' and 'object@meta.data':\n", 
      "adding gene set mean expression values in 'object@meta.data$gene.set.score'")
 se.dc.mus@meta.data$PAM_score <- mean.exp
}

# plot
png("~/t.gondi_dc/res/GEO_comparison/PAM_score_UMAP.png", width = 15, height = 10, units = "cm", res = 600)
FeaturePlot(object = se.dc.mus, 
            features = "PAM_score", pt.size = 2) +
              theme(axis.ticks = element_blank(),
                              axis.text = element_blank(), 
                              axis.title = element_blank(), 
                    legend.position = "right") +
               scale_color_gradientn("mean expression", colors = c("darkblue", "mediumblue", "yellow3", "yellow2","yellow1", "yellow")) 
  
dev.off()


```


Gene set expression across conditions (averaged over cells) can also be visualized in a heatmap

```{r, heatmap function}

#make annotation_col
annotation_col <- data.frame(CellType = factor(rep(c("GM_DC", "GM_Mac"), each = 8)))
rownames(annotation_col) <- colorder
colorder <- c("ctrl_GM_DC", "PTG_Lys3h_GM_DC" , "PTG_3h_GM_DC", "PTG_12h_GM_DC", "LDM_Lys3h_GM_DC", "LDM_3h_GM_DC" , "LDM_12h_GM_DC", "LPS_GM_DC", "ctrl_GM_Mac", "PTG_Lys3h_GM_Mac" , "PTG_3h_GM_Mac", "PTG_12h_GM_Mac", "LDM_Lys3h_GM_Mac", "LDM_3h_GM_Mac" , "LDM_12h_GM_Mac", "LPS_GM_Mac")
#add celltypes to the data in a complexheatmap
annotation_col$Strain <- paste(c("Control", "PTG", "PTG", "PTG", "LDM", "LDM", "LDM", "LPS", "Control", "PTG", "PTG", "PTG", "LDM", "LDM", "LDM", "LPS"))
annotation_col$Time <- as.integer(paste(c(3,3,3,12,3, 3, 12, 3, 3, 3, 3, 12, 3, 3,12, 3)))



#create dataframe with colors 

#data refers to a sc expression matrix you want to perform the analysis on 
#meta refers to the data you want to group your data by (e.g. clusters, celltypes, etc.)
#cond refers to the conditions you would like to display in the heatmap (numbers 1:16) (default is all conditions)
#seed refers to the features you would like to include (charactervector of genes to display)
#ann refers to the annotations you would like to have displayed in the heatmap  (default is all 3 annotations)

mat <- t(as.matrix(se.dc.mus@assays$SCT@data))
ann_cols <- list(CellType = c("GM_DC" = "mediumorchid4", "GM_Mac" = "turquoise4"), Strain = c("PTG" = "darkorange3", "LDM" = "blue3", "Control" = "forestgreen", "LPS" = "red3"), Time = c("gray20", "gray87"))

hmat_o <- hmat_mousetoxo(data = mat, meta = se.dc.mus$cell_class_type)

#make the list shorter and ensure the genes shown in the GO analysis are highlighted 

#for main figure: heatmap lps, heatmap cpGBs - add the remaining ones in the supplement
pdf("~/t.gondi_dc/res/figures/fig2/heatmap_lps.pdf")
phmat_original(hmat = hmat_o,
               cond = c(1:16), 
               seed = LPS_unique)
dev.off()

pdf("~/t.gondi_dc/res/figures/fig2/heatmap_cpGB.pdf")
phmat_original(hmat = hmat_o,
               cond = c(1:16), 
               seed = cpGB_unique)
dev.off()


```


```{r, compare data}

se.dc.mus <- SetIdent(se.dc.mus, value = "cell_class")
cellclass.markers <- FindAllMarkers(se.dc.mus, logfc.threshold = 0.5)
se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")
#identify differential expressed genes for each original condition 
LDM_12h_markers <- subset(cellclass.markers, cellclass.markers$cluster == "LDM_12h")
LDM_12h_markers <- subset(LDM_12h_markers$gene, LDM_12h_markers$p_val_adj < 0.05)
LDM_3h_markers <- subset(cellclass.markers, cellclass.markers$cluster == "LDM_3h")
LDM_3h_markers <- subset(LDM_3h_markers$gene, LDM_3h_markers$p_val_adj < 0.05)

PTG_12h_markers <- subset(cellclass.markers, cellclass.markers$cluster == "PTG_12h")
PTG_12h_markers <- subset(PTG_12h_markers$gene,  PTG_12h_markers$p_val_adj < 0.05)
PTG_3h_markers <- subset(cellclass.markers, cellclass.markers$cluster == "PTG_3h")
PTG_3h_markers <- subset(PTG_3h_markers$gene, PTG_3h_markers$p_val_adj < 0.05)



#Look at intersection differential expressed genes 

#make list for each toxo-condition and plot them in a table 
prr_list <- list(gene.set_LPS, gene.set_PAM, gene.set_cpGB, gene.set_HMW, gene.set_cGAMP, gene.set_Sev, gene.set_zymosan)
prr_unique <- set_names(list(LPS_unique, Pam_unique, cpGB_unique, HMW_unique, cGAMP_unique, Sev_unique, zymosan_unique), nm = c("LPS", "prr", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan"))


PTG_12h_int <- set_names(sapply(prr_unique, function(x){
  intersect(PTG_12h_markers, x)
}), nm = c("LPS", "PAM", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan"))

PTG_3h_int <- set_names(sapply(prr_unique, function(x){
  intersect(PTG_3h_markers, x)
}), nm = c("LPS", "PAM", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan"))

LDM_12h_int <- set_names(sapply(prr_unique, function(x){
  intersect(LDM_12h_markers, x)
}), nm = c("LPS", "PAM", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan"))

LDM_3h_int <- set_names(sapply(prr_unique, function(x){
  intersect(LDM_3h_markers, x)
}), nm = c("LPS", "PAM", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan"))


PTG_12h_int
PTG_12h_int
PTG_3h_int
LDM_12h_int
LDM_3h_int

overlap <- data.frame(PTG_3h = sapply(PTG_3h_int, function(x){length(x)}), PTG_12h = sapply(PTG_12h_int, function(x){length(x)}), LDM_3h = sapply(LDM_3h_int, function(x){length(x)}), LDM_12h = sapply(LDM_12h_int, function(x){length(x)}) )

overlap$PRR <- rownames(overlap)
overlap <- gather(overlap, "condition", value = "nGenes", 1:4)
overlap


update_geom_defaults("point",list(size=3))

#Export figure for figure 2
ggplot(overlap, aes(PRR, nGenes, color = condition)) + 
  geom_point(position = position_dodge(width=0.3), na.rm = T) +
  scale_color_manual(values = cell.cols[c(2,3,6,7)]) + 
  theme_classic() 
  
```

```{r, create a binary heatmap to look at gene expression in each condition, fig.height= 5, fig.width=5 }

#create an empy dataframe 
prr_list <- c(gene.set_LPS, gene.set_PAM, gene.set_cpGB, gene.set_HMW, gene.set_cGAMP, gene.set_Sev, gene.set_zymosan)
prr_unique <- c(LPS_unique, Pam_unique, cpGB_unique, HMW_unique, cGAMP_unique, Sev_unique, zymosan_unique)

#intersect prr with the toxodata 
PTG_12h_int <- intersect(PTG_12h_markers, prr_unique)
PTG_3h_int <- intersect(PTG_3h_markers, prr_unique)
LDM_12h_int <- intersect(LDM_12h_markers, prr_unique)
LDM_3h_int <- intersect(LDM_3h_markers, prr_unique)

toxo_int <- union(LDM_3h_int, LDM_12h_int)
toxo_int <- union(toxo_int,PTG_3h_int)  
toxo_int <- union(toxo_int, PTG_12h_int)

genes <- union(prr_unique, toxo_int)

bin.hmat <- data.frame(matrix(ncol = 11, nrow = length(genes)))

colnames(bin.hmat) <- c("LPS", "PAM", "cpGB", "HMW", "cGAMP", "SeV", "Zymosan", "LDM_3h", "LDM_12h", "PTG_3h", "PTG_12h")
rownames(bin.hmat) <- genes

bin.hmat[1:5,1:5]

#fill empty dataframe with values of each gene iterating over rows and columns 


bin.hmat[,1] <- ifelse(rownames(bin.hmat) %in% gene.set_LPS, 1, 0)
bin.hmat[,2] <- ifelse(rownames(bin.hmat) %in% gene.set_PAM, 1, 0)
bin.hmat[,3] <- ifelse(rownames(bin.hmat) %in% gene.set_cpGB,1, 0)
bin.hmat[,4] <- ifelse(rownames(bin.hmat) %in% gene.set_HMW, 1, 0)
bin.hmat[,5] <- ifelse(rownames(bin.hmat) %in% gene.set_cGAMP, 1, 0)
bin.hmat[,6] <- ifelse(rownames(bin.hmat) %in% gene.set_Sev, 1, 0)
bin.hmat[,7] <- ifelse(rownames(bin.hmat) %in% gene.set_zymosan, 1, 0)
bin.hmat[,8] <- ifelse(rownames(bin.hmat) %in% LDM_3h_markers, 1, 0)
bin.hmat[,9] <- ifelse(rownames(bin.hmat) %in% LDM_12h_markers, 1, 0)
bin.hmat[,10] <- ifelse(rownames(bin.hmat) %in% PTG_3h_markers, 1, 0)
bin.hmat[,11] <- ifelse(rownames(bin.hmat) %in% PTG_12h_markers, 1, 0)


bin.hmat2 <- t(bin.hmat)
pheatmap(bin.hmat2, angle_col =  c(315), fontsize_col = 8, cellheight = 10, cellwidth = 10, cluster_cols = F, color = colorRampPalette(c("#440154" ,"white", "#FDE725"))(200))

#show only genes present in toxoplasma infection

tox.overlap <- which(rowSums(bin.hmat[,c(8:11)]) >= 1)

pdf("~/t.gondi_dc/res/figures/fig2/overlap_genes_hmat.pdf")
pheatmap(bin.hmat[tox.overlap,], angle_col =  c(45), fontsize_col = 8,fontsize_row = 8, cluster_rows = F,
         cellheight = 10, cellwidth = 15, cluster_cols = F, color = c("#440154","#FDE725"))
dev.off()

```


We also would like to run an co-expression network 

```{r, load libraries, load and match data}

#load package
library(WGCNA)

#get preprocessed mouse matrix
M_mouse <- as.data.frame(se.dc.mus@assays$SCT@counts)
ncol(M_mouse)
#get pre-processed toxo matrix
se.toxo <- readRDS("~/t.gondi_dc/res/seurat-objects/toxo.object_150721.RDS")
M_toxo <- as.data.frame(se.toxo@assays$SCT@counts)
ncol(M_toxo)
#Filter columns of mouse data to only have columns containing toxo genes 

M_mouse <- dplyr::select(M_mouse, matches(colnames(M_toxo)))
ncol(M_mouse)

M_toxo <- as.matrix(dplyr::select(M_toxo, matches(colnames(M_mouse))))
ncol(M_toxo)

M_mouse <- as.matrix(M_mouse)

```

```{r, create correlation matrix}

set.seed(3)
corM = bicor(t(M_toxo),t(M_mouse))
corM[1:5,1:5]

hist(corM)

```

So we first find the top correlated interactors for each gene, and were only looking for interactions between moue and toxo.
We also only save 20 interactors for each gene - its unlikely anyone creates KNN graphs with a higher K. If they do there are much better tools than KNN graphs.

```{r, interactor defintion}

# Is it only selecting interactors with positive correlations here? 

TopIntGenes_mouse = matrix(rownames(M_toxo)[apply(FUN=order, X=corM, MARGIN=2, decreasing=T, na.last=T)[1:20,]],nrow=20)
TopIntGenes_toxo = matrix(rownames(M_mouse)[apply(FUN=order, X=corM, MARGIN=1, decreasing=T, na.last=T)[1:20,]],nrow=20)
colnames(TopIntGenes_mouse) = rownames(M_mouse)
colnames(TopIntGenes_toxo) = rownames(M_toxo)

```


```{r, correlation values}

TopIntVal_mouse = matrix(ncol=ncol(TopIntGenes_mouse),nrow=nrow(TopIntGenes_mouse))
for(ii in 1:ncol(TopIntVal_mouse)){TopIntVal_mouse[,ii] = corM[TopIntGenes_mouse[,ii],ii]}

TopIntVal_toxo = matrix(ncol=ncol(TopIntGenes_toxo),nrow=nrow(TopIntGenes_toxo))
for(ii in 1:ncol(TopIntVal_toxo)){TopIntVal_toxo[,ii] = corM[ii,TopIntGenes_toxo[,ii]]}

#combine matrices

TopIntGenes = cbind(TopIntGenes_mouse, TopIntGenes_toxo)
TopIntVal = cbind(TopIntVal_mouse, TopIntVal_toxo)
colnames(TopIntVal) = colnames(TopIntGenes)


```

For the resource, we have now pre-calculated all the objects we need. The rest will be performed in real time based on user input.

The user will select the following parameters:
seed = the seed genes we build networks from
K = number of nearest neighbors
Nsteps = Number of steps we take from the seed genes.
cutoffValue = the hard cutoff value

```{r, create function for plotting, fig.width = 18, fig.height = 15}

library(igraph)


f.KNN_mousetoxo = function(seed, K=3, Nsteps=2, cutoffValue=0.65){

# We first filter down the Top Interactors tables to the chosen number of interactors.
TopIntGenes2 = TopIntGenes[1:K,]
TopIntVal2 = TopIntVal[1:K,]

# We then remove interactors with a weight below 0.65.
if(cutoffValue<0){cutoffValue = 0; print("Warning: CutoffValue forced to 0!")}
x.remove = abs(TopIntVal2) < cutoffValue
TopIntGenes3 = TopIntGenes2; TopIntGenes3[x.remove] = NA
TopIntVal3 = TopIntVal2; TopIntVal3[x.remove] = NA

# Now we pick out the nearest neighbors, Nsteps steps away.
seed2 = seed
if(Nsteps>1){for(ii in 1:(Nsteps-1)){
  seed2 = c(seed2,TopIntGenes3[,seed2])
  seed2 = unique(seed2[!is.na(seed2)])
}}
TopIntGenes4 = TopIntGenes3[,seed2]
TopIntVal4 = TopIntVal3[,seed2]

# These objects are now transformed to an edgelist:
keep.x = !is.na(as.vector(TopIntGenes4))
elist1 = rep(colnames(TopIntGenes4), each=nrow(TopIntGenes4))[keep.x]
elist2 = as.vector(TopIntGenes4)[keep.x]
weights = as.vector(TopIntVal4)[keep.x]

# There may be different fancy packages to create interactive plots, but I will use a 
# straightforward solution: igraph

gr.x = graph.data.frame(cbind(elist1,elist2), directed = T)

# You can plot this graph, but it doesn't look very good. There are many things we can do to improve it. 
V(gr.x)$label.family = "sans"
V(gr.x)$label.cex = c(0.7)
V(gr.x)$label.font = c(2)
V(gr.x)$label.color = c("#333333")
#V(gr.x)$label.dist = c(-0.5)
#V(gr.x)$label.degree = pi/6
E(gr.x)$arrow.size = 0.5
E(gr.x)$width = abs(weights)*2
V(gr.x)$name = make.names(V(gr.x)$name)
#V(gr.x)$color[V(gr.x)$name %in% make.names(PTG_12h_markers)] = "#66CCFF"
#V(gr.x)$color[V(gr.x)$name %in% make.names(LDM_12h_markers)] = "yellow3"
V(gr.x)$color[V(gr.x)$name %in% make.names(colnames(TopIntGenes_mouse))] = "#66CCFF"
V(gr.x)$color[V(gr.x)$name %in% make.names(colnames(TopIntGenes_toxo))] = "yellow3"
V(gr.x)$frame.color[V(gr.x)$name %in% make.names(colnames(TopIntGenes_mouse))] = "#3399CC"
V(gr.x)$frame.color[V(gr.x)$name %in% make.names(colnames(TopIntGenes_toxo))] = "yellow4"
#V(gr.x)$shape[V(gr.x)$name %in% make.names(colnames(TopIntGenes_mouse))] = "circle"
#V(gr.x)$shape[V(gr.x)$name %in% make.names(colnames(TopIntGenes_toxo))] = "circle"

# We set the vertex sizes according to (minimum) distance from seed genes
dist.x = distances(gr.x)
for(ii in 1:ncol(dist.x)){dist.x[ii,ii] = NA}
dist.x = dist.x[rownames(dist.x)%in%seed,]
dist.x2 = apply(FUN=min, X=dist.x, MARGIN=2,na.rm=T)
V(gr.x)$size = 2+10*(2^-dist.x2[V(gr.x)$name])


# Color for interactions with positive correlations is set to red. 
# Negative, if such exist, are colored blue.
E(gr.x)$color[weights>0] = "#CC3333"
E(gr.x)$color[weights<0] = "blue"

# We can also map information to colors and shape of vertices.
# There are two obvious options: 
# 1) Different shapes for mouse and toxo, node colors according to a selected cluster from the umap.
# 2) Make each node a pie chart with one piece for each cluster. Then distinction between mouse
#    and toxo will have to be made some other way.
#
# I will leave this choice to a future discussion. For now, here is a simple function to translate
# numeric values into a color scale, like a typical heatmap.

f.TranslateToColor = function(val){
  bins.x = seq(from=-max(abs(val),na.rm=T)-0.01, to=max(abs(val),na.rm=T)+0.01, length.out=101)
  cval = c()
  for(ii in 1:length(val)){cval[ii] = sum(bins.x<val[ii])}
  cvec = colorRampPalette(c("blue", "dodgerblue", "white", "orange", "red"))(100)[cval]
  if(is.matrix(val)){
    cvec = matrix(cvec,nrow=nrow(val))
    rownames(cvec) = rownames(val)
    colnames(cvec) = colnames(val)
  } else {
    names(cvec) = names(val)
  }
  return(cvec)
}
par(mar=c(0,0,0,0)+.1)
plot(gr.x, asp = 0) 
return(gr.x)


}



pdf("~/t.gondi_dc/res/figures/fig3/PTG_markers_network.pdf")
gr.1 <- f.KNN_mousetoxo(c(PTG_12h_markers), cutoffValue = 0.5, Nsteps = 2, t.col ="maroon4")
dev.off()

pdf("~/t.gondi_dc/res/figures/fig3/LDM_markers_network.pdf")
gr.2 <- f.KNN_mousetoxo(c(LDM_12h_markers), cutoffValue = 0.5, Nsteps = 2, t.col = "darkgreen")
dev.off()

gr.3 <- f.KNN_mousetoxo(c(PTG_12h_markers), cutoffValue = 0.5, Nsteps = 100)

```

```{r, run subgraph of graph in new function}


#gr = graphobject
#goi = gene of interest

subgraph_mtoxo = function(gr, goi){
mem.x = clusters(gr)$membership
index.x = mem.x[V(gr)$name == goi]
gr.x2 = induced_subgraph(gr, which(mem.x==index.x))
par(mar=c(0,0,0,0)+.1)
plot(gr.x2, asp = 0) 
return(gr.x2)
}

subgraph_mtoxo(gr = gr.1, goi = "TGME49.278090")

```

```{r, make table for toxo and mouse genes from interaction graph}


go.table.tg$ensembl_gene_id <- gsub("-", ".",go.table.tg$ensembl_gene_id)
#gr = graph
#ann = annotation table - default is the BioMart table used for annotations 
get.toxo.table <- function(gr, ann = go.table.tg){
inf <- data.frame(clusters(gr)$membership)
colnames(inf) <- c("interaction_cluster")
inf$ensembl_gene_id <- rownames(inf)
int.table <- ann[ann$ensembl_gene_id %in% rownames(inf),]
tg.int <- dplyr::left_join(int.table,inf, by = "ensembl_gene_id")
return(tg.int)
}

tab1 <- get.toxo.table(gr.1)

write.table(tab1, "~/t.gondi_dc/res/interaction_table_toxo.csv", sep = ";", quote = F, row.names = F)

##Do the same for mouse genes

get.mouse.table <- function(gr, ann = go.table){
inf <- data.frame(clusters(gr)$membership)
colnames(inf) <- c("interaction_cluster")
inf$gene_symbol <- rownames(inf)
int.table <- ann[ann$gene_symbol %in% rownames(inf),]
mouse.int <- dplyr::left_join(int.table,inf, by = "gene_symbol")
return(mouse.int)
}

tab <- get.mouse.table(gr.3)

write.table(tab, "~/t.gondi_dc/res/interaction_table_mus.csv", sep = "\t", quote = F, row.names = F)

#gprofiler - enrichment
mouse.int.list <- names(clusters(gr.1)$membership[V(gr.1)$name %in% make.names(colnames(TopIntGenes_mouse))])
mus.int <- gost(mouse.int.list, organism = "mmusculus", sources = "GO:BP")

  print(ggplot(head(mus.int$result), aes(reorder(term_name, -log10(p_value)), -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "darkblue") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 20, name = "Reds")) + 
  scale_y_continuous() +
  xlab("Biological Process")+
  ylab("Enrichment Score") +
  theme(axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.background = element_blank()))

```

```{r, look at enrichment for individual interaction clusters}

#get unique genes for each interaction cluster
tab.n <- data.frame(gene = tab$gene_symbol, interaction.c = tab$interaction_cluster)
tab.n <- unique(tab.n)
tab.n <- tab.n[order(tab.n$interaction.c),]
min <- min(tab.n$interaction.c)
max <- max(tab.n$interaction.c)
clust.sizes <- setNames(sapply(c(min:max), function (x){
  length(which(tab.n$interaction.c %in% x))
}), nm =c(min:max))
#run cluster-wise gene ontologies
max(tab.n$interaction.c)
min(tab.n$interaction.c)

golysis <- function(data, cluster){
  dat <- list(data[data[,2] == cluster,][,1])
  res <- gost(query = dat, organism = "mmusculus", sources = "GO:BP", significant = F)
  
  g <- ggplot(head(res$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "darkblue") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 20, name = "Reds")) + 
  scale_y_continuous() +
  xlab("Biological Process")+
  ylab("Enrichment Score") +
  theme(axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.background = element_blank())
  return(g)
}

golysis(tab.n, 18)


```

```{r, alternative heatmaps}

heatmap.F(p_all[PTG_LDM_markers$gene, c(6:7)], clustermethod = "complete")

#ask about heatmap 

```


