---
title: "DC-analysis-new"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


PTG (ME49)  = type II (considered to be avirulent) with high expression/levels of IL-12
LDM (RH/GT1) = type I (considered virulent) 

```{r, load required packages, warning=FALSE}

library(Seurat)
library(dplyr)
library(Matrix)
library(gtable)
library(grid)
library(gridExtra)
library(rlang)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(ggplot2)
library(biomaRt)
library(tidyr)
library(plotly)
library(pheatmap)
library(forcats)
library(patchwork)
library(sctransform)
library(openxlsx)
suppressMessages(require(biomaRt))
library(EnhancedVolcano)
library(pheatmap)
```

Import the count matrix of the single cell sequencing experiment and subset it, so it only contains the counts for the mouse genome

```{r, import expression data and subset it}

#import expression matrix first 

dc.data <- read.csv("~/t.gondi_dc/data/cnt/cnt_mus_toxo_raw.csv", row.names = 1, header = T, check.names = FALSE)

#subset for only mouse genes: 
mus.genes <- rownames(dc.data)[grep("ENSMUSG",rownames(dc.data))]

head(mus.genes)
dc.data.mus <- dc.data[mus.genes,]

```

Convert ensembl-Ids to gene symbols in the matrix

```{r make gene symbol annotation file}

#use ensembl mart 

mart <- useMart("ensembl")

mart <- useDataset("mmusculus_gene_ensembl", mart = mart)

genes.table <- getBM(filters = "ensembl_gene_id", attributes = c("ensembl_gene_id", "external_gene_name", "description", "gene_biotype", "entrezgene_id"), values = rownames(dc.data.mus), mart = mart )

head(genes.table)

#write.table(genes.table, file = "gene_name_translation.tab")

head(genes.table)


#Give the data new unique rownames
genes.table$unique_gene_name <- NA
genes.table$unique_gene_name[!is.na(genes.table$external_gene_name)] <- make.unique(genes.table$external_gene_name[!is.na(genes.table$external_gene_name)])
genes.table$unique_id[!is.na(genes.table$ensembl_gene_id)] <- make.unique(genes.table$ensembl_gene_id[!is.na(genes.table$ensembl_gene_id)])

rownames(genes.table) <- genes.table$unique_id

#merge ensembl name and gene ID
dc.data.mus <- dc.data.mus[intersect(rownames(dc.data.mus), rownames(genes.table)), ]
newnames <- genes.table[rownames(dc.data.mus), ]$unique_gene_name

rownames(dc.data.mus) <- newnames

```

Create the seurat object for analysis - set 2 markers for "mature" and "immature" cell types of BDMCs to get an impression of the proportions of these two types for each condition in the data-set. However I would like to use the mature/immature classification we have calculated - find a way to use this instead if comparing numbers 

```{r, Create Seuratobject}

meta1 <- read.table("~/t.gondi_dc/data/meta/metadata-200430.csv", sep = ",", header = T, row.names = 1)

se.dc.mus <- CreateSeuratObject(dc.data.mus, min.cells = 3, min.features = 200, project = "DC-TG-FH-M", meta.data = meta1)

se.dc.mus[["percent.mt"]] <- PercentageFeatureSet(se.dc.mus,pattern = "mt-")
se.dc.mus[["percent.mature"]] <- PercentageFeatureSet(se.dc.mus, features = "mature")
se.dc.mus[["percent.immature"]] <- PercentageFeatureSet(se.dc.mus, features= "immature")

grayscale = c("gray5","gray15", "gray30", "gray55", "gray70", "gray85", "gray95", "gray100")

VlnPlot(se.dc.mus, features = c("percent.mt","percent.mature", "percent.immature"), ncol = 2, group.by = "cell_class", cols = grayscale)

#subset the cells with a high percentage of mitochondrial genes 
se.dc.mus <- subset(se.dc.mus, subset = percent.mt < 10)

#subset everything but the LPS cells
se.dc.mus.f <- subset(se.dc.mus, subset = cell_class %in% c("ctrl", "LDM_12h", "LDM_3h", "PTG_12h", "PTG_3h", "LDM_Lys3h", "PTG_Lys3h"))

#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/qc/RNA-andFeatureCount-DC.png", width = 35, height = 15, res = 300, units = "cm") #Units are pixels by default
print(VlnPlot(se.dc.mus, features = c("nCount_RNA", "nFeature_RNA"), ncol = 2, group.by = "cell_class", cols = grayscale))
#dev.off()

```

```{r, normalize the data, results="hide"}

#Run in console 
se.dc.mus <- SCTransform(se.dc.mus)

```

```{r, Run dimensionality reduction and clustering}

se.dc.mus <- RunPCA(se.dc.mus)

print(se.dc.mus[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(se.dc.mus, dims = 1:2, reduction = "pca")

DimPlot(se.dc.mus,  reduction = "pca", group.by = "cell_class", cols = "Dark2")

DimHeatmap(se.dc.mus, dims = 1:15, cells = 500, balanced = TRUE)

ElbowPlot(se.dc.mus, ndims= 20, reduction ="pca")

se.dc.mus <- FindNeighbors(se.dc.mus, dims = 1:15)
se.dc.mus <- FindClusters(se.dc.mus) #start with default resolution of 0.8
#inspect the clusters
head(Idents(se.dc.mus), 5)

se.dc.mus <- RunUMAP(se.dc.mus, dims = 1:10, spread = 0.5, min.dist = 0.1)
#min.dist let's me play around with the thightnes of the cells within the clusters
#the spread let's me play around with the coordinates of the distance between the clusters 
##default settings are -> min.dist: 0.3, spread: 1

DimPlot(se.dc.mus, reduction = "umap", group.by = "seurat_clusters") 

DimPlot(se.dc.mus, reduction = "umap", group.by = "cell_class")


saveRDS(se.dc.mus, "~/t.gondi_dc/res/seurat-objects/seurat-dc-2021-01-04")

```

Plot the Umap and visualize the clusters as well as the original cell identity

```{r, umap coordinates and plots}

c.cols <- c("deeppink3", "darkcyan", "darkgoldenrod3", "chartreuse4","darkorange", "brown3", "deepskyblue3", "blueviolet", "firebrick")
cell.cols <- c("forestgreen", "blue3","dodgerblue3", "olivedrab3","red4", "magenta4", "deeppink3","seagreen3")
umap.dc.m.clusters <- DimPlot(se.dc.mus, reduction = "umap", group.by = "seurat_clusters", cols = c.cols)
umap.dc.m.cellclass <- DimPlot(se.dc.mus, reduction = "umap", group.by = "cell_class", cols = cell.cols)

#extract the umap coordinates for possible croos-references
umap.coordinates.dc <- as.data.frame(Embeddings(se.dc.mus, reduction = "umap"))


#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/DC-final-umap.png", width = 35, height = 15, res = 300, units = "cm") #Units are pixels by default
print(cowplot::plot_grid(umap.dc.m.cellclass, umap.dc.m.clusters))
#dev.off()


```


```{r, calculate percentages of each condition within individual clusters}

#write sapply function to give vector for each cluster and condition and create a dataframe with the information 
clusters <- c(0:8)

LPS_perCluster <- sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LPS" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_3h_perCluster <- sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_12h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_12h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_12h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_12h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

ctrl_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "ctrl" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

LDM_Lys3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "LDM_Lys3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

PTG_Lys3h_perCluster <-  sapply(X = clusters, function(cluster) {
  length(which(se.dc.mus@meta.data$cell_class == "PTG_Lys3h" & se.dc.mus@meta.data$seurat_clusters == cluster))
})

n1 <- c("cluster0", "cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8")
identpercluster <- data.frame(n1, LDM_3h_perCluster, PTG_3h_perCluster, LDM_12h_perCluster, PTG_12h_perCluster, ctrl_perCluster, LDM_Lys3h_perCluster, PTG_Lys3h_perCluster, LPS_perCluster)
colnames(identpercluster) <- c("cluster", "LDM 3h", "PTG 3h", "LDM 12h", "PTG 12h", "uninfected", "LDM Lysate", "PTG Lysate", "LPS")
identpercluster <- gather(identpercluster, key = "identity", value = "cellcount", -cluster)

#calculate percentages using an apply function 
percent <- function(x){
  ((x/sum(identpercluster$cellcount))*100)
}

identpercluster$percent <- identpercluster[,3]/(sum(identpercluster$cellcount))*100

#Plot results:
#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/originialIDpercluster_percent.png", width = 20, height = 10, unit = "cm", res = 300)
print(ggplot(identpercluster, aes(cluster, percent, fill= identity)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("") +
  ylab ("relative proportion of cells") + 
  scale_fill_manual(values = c( "#D95F02", "#7570B3" ,"#E7298A", "#66A61E" ,"#E6AB02" ,"#A6761D", "#666666" ,"#1B9E77")) + 
  theme_classic() + 
  theme(legend.title = element_blank()))
#dev.off()

```


```{r, scale the calcualted percentages}
#calculate percentages of individual clusters with normalizing to the individual cluster 

c0.p <- subset(identpercluster, cluster == "cluster0")
c1.p <- subset(identpercluster, cluster == "cluster1")
c2.p <- subset(identpercluster, cluster == "cluster2")
c3.p <- subset(identpercluster, cluster == "cluster3")
c4.p <- subset(identpercluster, cluster == "cluster4")
c5.p <- subset(identpercluster, cluster == "cluster5")
c6.p <- subset(identpercluster, cluster == "cluster6")
c7.p <- subset(identpercluster, cluster == "cluster7")
c8.p <- subset(identpercluster, cluster == "cluster8")

cp.l <- c(c0.p$cellcount, c1.p$cellcount, c2.p$cellcount, c3.p$cellcount,c4.p$cellcount, c5.p$cellcount, c6.p$cellcount, c7.p$cellcount , c8.p$cellcount)


c0.p$percent.cluster <-  c0.p$cellcount/sum(c0.p$cellcount)*100
c1.p$percent.cluster <-  c1.p$cellcount/sum(c1.p$cellcount)*100
c2.p$percent.cluster <-  c2.p$cellcount/sum(c2.p$cellcount)*100
c3.p$percent.cluster <-  c3.p$cellcount/sum(c3.p$cellcount)*100
c4.p$percent.cluster <-  c4.p$cellcount/sum(c4.p$cellcount)*100
c5.p$percent.cluster <-  c5.p$cellcount/sum(c5.p$cellcount)*100
c6.p$percent.cluster <-  c6.p$cellcount/sum(c6.p$cellcount)*100
c7.p$percent.cluster <-  c7.p$cellcount/sum(c7.p$cellcount)*100
c8.p$percent.cluster <-  c8.p$cellcount/sum(c8.p$cellcount)*100

identpercluster1 <- dplyr::bind_rows(c0.p,c1.p, c2.p,c3.p, c4.p, c5.p, c6.p, c7.p, c8.p)

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/cellIDpercluster-norm.png", width = 20, height = 10, units = "cm", res = 300) 
  print(ggplot(identpercluster1, aes(cluster, percent.cluster, fill= identity)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("") +
  ylab ("cells of original condition per cluster [%]") + 
  scale_fill_manual(values = c( "#D95F02", "#7570B3" ,"#E7298A", "#66A61E" ,"#E6AB02" ,"#A6761D", "#666666" ,"#1B9E77")) + 
  theme_classic() + 
  theme(legend.title = element_blank()))
#  dev.off()

```

Next we perform a variety of differential gene expression analysis based on multiple factors and conditions

1. Find markers based on the original cell class in the metadata, i.e. based on the experimental condition

```{r, Find markers for original identities}

se.dc.mus <- SetIdent(se.dc.mus, value = "cell_class")

dc.markers.original <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.5)

dc.markers.original <- subset(dc.markers.original, p_val_adj < 0.05)

#write.xlsx(dc.markers.original, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-orginal-IDs.xlsx")


```

2. Find markers for the identified clusters in the cluster analysis 

```{r, Find Markers according to cluster identity and subset them for later gprofiler analysis}

se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")

#Find markers for all clusters
se.dc.markers <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.5)

#subset them for only significantly differential expressed genes 
se.dc.markers <- subset(se.dc.markers, p_val_adj < 0.05)

#look at differential expressed genes between the two clusters that comprise most of the 12h time-point (independent of strain, cluster 3 and cluster 5)
dc.12h.markers <- FindMarkers(se.dc.mus, ident.1 = 4, ident.2 = 3, min.pct = 0.25, logfc.threshold = 0.5)
dc.12h.markers <- subset(dc.12h.markers, p_val_adj < 0.05)
#visualize in a volcano plot
volcano12h.3.4 <- EnhancedVolcano(dc.12h.markers , lab = rownames(dc.12h.markers),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcano-12hDEG-leftPTG-rightLDM.png", width = 15, height = 25, units = "cm", res= 300)
print(volcano12h.3.4)
#dev.off()

#write.xlsx(dc.12h.markers,"~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/DEG-cluster3-cluster5.xlsx", row.names = T)

```


```{r, dcreate a phylogenetic tree of the conidtions based on the expression matrix, width = 20, height =10}

clust.tree <- BuildClusterTree(se.dc.mus, assay = "SCT", slot = "data")
clust.tree <- Tool(clust.tree, slot = "BuildClusterTree")
ggtree(clust.tree, layout = "roundrect") + 
  geom_tiplab() +
  geom_hilight(node = 7, fill = "seagreen3", alpha =.6) +
  geom_hilight(node = 4, fill = "olivedrab3" , alpha =.6) +
  geom_hilight(node = 3, fill = "forestgreen" , alpha =.6) +
  geom_hilight(node = 1, fill = "dodgerblue" , alpha =.6) +
  geom_hilight(node = 8, fill = "red3" , alpha =.6) +
  geom_hilight(node = 2, fill = "blue3" , alpha =.6) +
  geom_hilight(node = 6, fill = "magenta4" , alpha =.6) +
  geom_hilight(node = 5, fill = "deeppink3" , alpha =.6)

```


```{r, run GO term GO term analyis based on clusters}

se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")

dc.markers <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.5)

dc.markers <- subset(dc.markers, p_val_adj < 0.05)

dc.markers.C0 <- subset(dc.markers.original, dc.markers$cluster == 0)
dc.markers.C1 <- subset(dc.markers.original, dc.markers$cluster == 1)
dc.markers.C2 <- subset(dc.markers.original, dc.markers$cluster == 2)
dc.markers.C3 <- subset(dc.markers.original, dc.markers$cluster == 3)
dc.markers.C4 <- subset(dc.markers.original, dc.markers$cluster == 4)
dc.markers.C5 <- subset(dc.markers.original, dc.markers$cluster == 5)
dc.markers.C6 <- subset(dc.markers.original, dc.markers$cluster == 6)
dc.markers.C7 <- subset(dc.markers.original, dc.markers$cluster == 7)
dc.markers.C8 <- subset(dc.markers.original, dc.markers$cluster == 8)


marker.list <- list(dc.markers.C0, dc.markers.C1, dc.markers.C2, dc.markers.C3, dc.markers.C4, dc.markers.C4, dc.markers.C5, dc.markers.C6, dc.markers.C7, dc.markers.C8)

dc.markers.n <- c("dc.markers.C0","dc.markers.C1", "dc.markers.C2", "dc.markers.C3","dc.markers.C4", "dc.markers.C5", "dc.markers.C6", "dc.markers.C7", "dc.markers.C8")

subset.markers.l <- setNames(lapply(marker.list, function(x) {
  subset(x, avg_logFC > 0.5 & p_val_adj < 0.01)
}), nm = dc.markers.n)

```


```{r, run GO term analysis with different databases}

library(gprofiler2)

gostres.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x$gene, organism = "mmusculus")
}), nm = c("gostres.C0", "gostres.C1", "gostres.C2", "gostres.C3", "gostres.C4", "gostres.C5", "gostres.C6","gostres.C7","gostres.C8"))

#subset for GO:BP only
gostres.BP.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x$gene, organism = "mmusculus", sources = "GO:BP")
}), nm = c("gostres.BP.C0", "gostres.BP.C1", "gostres.BP.C2", "gostres.BP.C3", "gostres.BP.C4", "gostres.BP.C5", "gostres.BP.C6", "gostres.BP.C7", "gostres.BP.C8"))

#subset for Kegg only

gostres.KEGG.l <- setNames(lapply(subset.markers.l, function(x){
  gost(query = x$gene, organism = "mmusculus", sources = "KEGG")
}), nm = c("gostres.KEGG.C0", "gostres.KEGG.C1", "gostres.KEGG.C2", "gostres.KEGG.C3", "gostres.KEGG.C4", "gostres.KEGG.C5", "gostres.KEGG.C6", "gostres.KEGG.C7"))

#make a list 
cUmapcluster <- c("deeppink3", "darkcyan", "darkgoldenrod3", "chartreuse4","darkorange", "brown2", "deepskyblue3", "blueviolet", "firebrick")
#Plot the data

print(ggplot(head(gostres.BP.l$gostres.BP.C4$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "firebrick") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 20, name = "Reds")) + 
  scale_y_continuous() +
  xlab("Biological Process")+
  ylab("Enrichment Score") +
  theme(axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.background = element_blank()))



GO.all.1 <- cowplot::plot_grid(p.go.0,NULL,p.go.4,NULL,p.go.6, ncol= 5, nrow = 1, rel_widths = c(5,0.1,5,0.1,5))
GO.all.1

GO.all.2 <- cowplot::plot_grid(p.go.1,NULL,p.go.2, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.2
GO.all.3 <- cowplot::plot_grid(p.go.3,NULL,p.go.5, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.3
GO.all.4 <- cowplot::plot_grid(p.go.7,NULL,p.go.8, ncol= 3, nrow = 1, rel_widths = c(5,0.1,5))
GO.all.4

#Save the plots all together 
#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/GO-enrichment/GO-enrichments-clusterbased-4.png", width = 60, height = 10, res = 300, units = "cm") #Units are pixels by default
print(GO.all.4)
#dev.off()

#Save the plots individually

#png(filename = "~/Desktop/PhDprojectFranziskaHildebrandt/TG-DC-Analysis/Results-200507-FiguresAndTables/GO-enrichment/GO-enrichment-C0.png", width = 30, height = 15, res = 300, units = "cm") #Units are pixels by default
print(p0)
#dev.off()

```

In addition: Try to run the DGEA according to the original cell-identity and look at the results and the GO-term analysis in the DCs! 

```{r, DEGs according to original cell identity, heatmap, GOterms}

se.dc.mus <- SetIdent(se.dc.mus, value = "cell_class")

markers.cellident <- FindAllMarkers(se.dc.mus, min.pct = 0.25, logfc.threshold = 0.25)


markers.cellident %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)

top20.all.i <- markers.cellident %>% group_by(cluster) %>% top_n(n=20, wt =avg_logFC)
top.20.all.i <- DoHeatmap(se.dc.mus, features = top20.all.i$gene)
top.20.all.i

markers.control <- subset(markers.cellident, markers.cellident$cluster == "ctrl")
markers.LDM12h <- subset(markers.cellident, markers.cellident$cluster == "LDM_12h")
markers.LDM_3h <- subset(markers.cellident, markers.cellident$cluster == "LDM_3h")
markers.LDM_Lys3h <- subset(markers.cellident, markers.cellident$cluster == "LDM_Lys3h")
marker.LPS <- subset(markers.cellident, markers.cellident$cluster == "LPS")
markers.PTG_12h <- subset(markers.cellident, markers.cellident$cluster == "PTG_12h")
markers.PTG_3h <- subset(markers.cellident, markers.cellident$cluster == "PTG_3h")
markers.PTG_Lys3h <- subset(markers.cellident, markers.cellident$cluster == "PTG_Lys3h")


marker.list.cellident <- list(markers.control, markers.LDM12h,markers.LDM_3h, markers.LDM_Lys3h, markers.PTG_12h,markers.PTG_3h, markers.PTG_Lys3h, marker.LPS)
dc.markers.i.n <- c("ctrl", "LDM_12h","LDM_3h","LDM_Lys3h","PTG_12h","PTG_3h","PTG_Lys3h", "LPS")

subset.i.l <- setNames(lapply(marker.list.cellident, function(x) {
  subset(x, p_val_adj < 0.05)
}), nm = dc.markers.i.n)


###GO terms 

gostres.i.l <- setNames(lapply(subset.i.l, function(x){
  gost(query = x$gene, organism = "mmusculus")
}), nm = dc.markers.i.n)

#subset for GO:BP only
gostres.BP.i.l <- setNames(lapply(subset.i.l, function(x){
  gost(query = x$gene, organism = "mmusculus", sources = "GO:BP")
}), nm = dc.markers.i.n)

#subset for Kegg only

gostres.KEGG.i.l <- setNames(lapply(subset.i.l, function(x){
  gost(query = x$gene, organism = "mmusculus", sources = "KEGG")
}), nm = dc.markers.i.n)


#Plot the data
c( "#D95F02", "#7570B3" ,"#E7298A", "#66A61E" ,"#E6AB02" ,"#A6761D", "#666666" ,"#1B9E77")
##For BP we only get significant data for 12h timepoints!

p.go.LDM.12h <- ggplot(head(gostres.BP.i.l$LDM_12h$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value), fill = -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "#D95F02") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 9, name = "Reds")) + 
  xlab("Biological Process")+
  ylab("Enrichment Score")

p.go.PTG.12h <- ggplot(head(gostres.BP.i.l$PTG_12h$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value), fill = -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "#E6AB02") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 9, name = "Reds")) + 
  xlab("Biological Process")+
  ylab("Enrichment Score")

p.go.LPS <- ggplot(head(gostres.BP.i.l$LPS$result, 10), aes(reorder(term_name, -log10(p_value)), -log10(p_value), fill = -log10(p_value))) +
  geom_bar(stat = "identity", color = "black", fill = "#66A61E") +
  coord_flip() +
  #scale_fill_gradientn(name = "Enrichment Score",colours =   RColorBrewer::brewer.pal(n = 9, name = "Reds")) + 
  xlab("Biological Process")+
  ylab("Enrichment Score")


GO.sig.originalID <- cowplot::plot_grid(p.go.LDM.12h,NULL,p.go.PTG.12h,NULL,p.go.LPS,NULL, ncol= 6, nrow = 1, rel_widths = c(5,0.1,5, 0.1,5,0.1))

#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/GO-enrichment/GO-enrichment-cell-id-sig-all.png", width = 50, height = 10, res = 300, units = "cm") #Units are pixels by default
print(GO.sig.originalID)
#dev.off()

```

There is a distinction between mature and immature DCs [Shalek A., 2013](https://pubmed.ncbi.nlm.nih.gov/23685454/) or GM-DCs (= mature) and GM-Macs (= immature) [Helft J., 2015](https://pubmed.ncbi.nlm.nih.gov/26084029/). Therefore we want to see whether the identified sub-populations in our UMAP correspond to these classifications by using markers that correspond to these classifications.

```{r, plot the mature and immature markers on umaps, fig.width= 12, fig.height= 6}

#Way to identify cluster identity? 

matdc <- c("Ccr7", "Ccl22", "Serpinb9", "Cd83")
immatdc <- c("Tnf", "Il1a", "Cxcl10", "C5ar1")

p1 <- FeaturePlot(se.dc.mus, features = c("Ccr7"), cols = c("lightgray", "mistyrose", "darkred"))
p2 <- FeaturePlot(se.dc.mus, features = c("Cd83"), cols = c("lightgray", "mistyrose", "darkred"))
p3 <- FeaturePlot(se.dc.mus, features = c("C5ar1"), cols = c("lightgray", "lightblue", "darkblue"))
p4 <- FeaturePlot(se.dc.mus, features = c("Il1a"), cols = c("lightgray", "lightblue", "darkblue"))

#png(filename = "~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/mature_vs_immature_umap.png", width = 30, height = 15, res = 300, units = "cm") #Units are pixels by default
print(plot_grid(p1,p2,p3,p4, ncol = 2, nrow = 2))
#dev.off()


```

```{r, AddModuleScore based on maturity}


matdc <- c("Ccr7", "Ccl22", "Serpinb9", "Cd83")
immatdc <- c("Tnf", "Il1a", "Cxcl10", "C5ar1")

se.dc.mus <- AddModuleScore(se.dc.mus, features = matdc, assay = "SCT", name = "matureDC", seed = 1)

head( se.dc.mus[[]])

# ModuleScores added for each gene of each cell -> 
# write if statement that says if 3 our of 4 are positive = mature, if 2 out of 4 = positive = intermediate, if 1 out of 4 = positive = immature! 

maturedcs1 <- sapply((se.dc.mus@meta.data$matureDC1), function(x) {
  if (x > 0)  paste0(1) else paste0(0)
  })
maturedcs2 <- sapply((se.dc.mus@meta.data$matureDC2), function(x) {
  if (x > 0)  paste0(1) else paste0(0)
  })
maturedcs3 <- sapply((se.dc.mus@meta.data$matureDC3), function(x) {
  if (x > 0)  paste0(1) else paste0(0) 
  })

maturedcs4 <- sapply((se.dc.mus@meta.data$matureDC4), function(x) {
  if (x > 0)  paste0(1) else paste0(0) 
  })

maturityscores <- data.frame(maturedcs1, maturedcs2, maturedcs3, maturedcs4, stringsAsFactors = F) %>% 
  mutate_all(as.numeric) 
rownames(maturityscores) <- rownames(se.dc.mus@meta.data)
maturityscores <- transform(maturityscores, sum = rowSums(maturityscores))

#attach maturity as a column to the dataframe: 

maturityscores <- as.data.frame(sapply((maturityscores$sum), function(x) {
  if(x >= 3) paste0("mature") 
  else paste0("immature")
  }))

rownames(maturityscores) <- rownames(se.dc.mus@meta.data)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = maturityscores, col.name = "maturation_state")

DimPlot(se.dc.mus, group.by = "cell_class")
DimPlot(se.dc.mus, group.by = "maturation_state")

FeaturePlot(se.dc.mus, features = "Ccr7")

```
Next we want to infer on the DEG between the maturity state/immune cell type and the original cell class of the metadata i.e. the infection state

```{r, pairwise DGEA for mature vs immature}

se.dc.mus$cluster_and_cell_class <- paste0(se.dc.mus$SCT_snn_res.0.8, "_", se.dc.mus$cell_class)
se.dc.mus$cell_class_maturation <- paste0(se.dc.mus$cell_class, "_", se.dc.mus$maturation_state)
se.dc.mus$cluster_maturation <- paste0(se.dc.mus$SCT_snn_res.0.8, "_", se.dc.mus$maturation_state)

se.dc.mus <- SetIdent(se.dc.mus, value = "cell_class_maturation")

id1 <- "ctrl_mature"
id2 <- "ctrl_immature"
control_maturity_m <- FindMarkers(se.dc.mus,ident.1 = id1, ident.2 = id2, min.pct = 0.25)

id3 <- c("LDM_3h_mature", "PTG_3h_mature")
id4 <- c("LDM_3h_immature", "PTG_3h_immature")
infected_maturity_m_3h <- FindMarkers(se.dc.mus,ident.1 = id3, ident.2 = id4, min.pct = 0.25)

id5 <- c("LDM_12h_mature", "PTG_12h_mature")
id6 <- c("LDM_12h_immature", "PTG_12h_immature")

infected_maturity_m_12h <- FindMarkers(se.dc.mus,ident.1 = id5, ident.2 = id6, min.pct = 0.25)

infected_maturity_m_12h

```

```{r, volcano plots for visualization of DEG}

volcano.infected.3h <- EnhancedVolcano(infected_maturity_m_3h, lab = rownames(infected_maturity_m_3h),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.infected.12h <- EnhancedVolcano(infected_maturity_m_12h, lab = rownames(infected_maturity_m_12h),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.control <- EnhancedVolcano(control_maturity_m , lab = rownames(control_maturity_m),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.infected.3h
volcano.infected.12h
volcano.control

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/pairwise-DE-infection-control-maturation-all.png", width = 20, height = 40, units = "cm", res = 300)
print(cowplot::plot_grid(volcano.control, volcano.infected.3h, infected.12h, nrow = 3, ncol = 1))
#dev.off()

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-3h-mature-vs-immature.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.infected.3h)
#dev.off()

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-12h-mature-vs-immature.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.infected.12h)
#dev.off()

#png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-control-mature-vs-immature.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.control)
#dev.off()
```

```{r, run the DGEA with the same trhesholds as used for volcano plots}

#Filter the datasets according to the same thresholds as in volcano plots: 

cmm <- FindMarkers(se.dc.mus,ident.1 = id1, ident.2 = id2, min.pct = 0.25, logfc.threshold = 1)
cmm <- subset(cmm, p_val_adj < 0.01)

imm3 <- FindMarkers(se.dc.mus,ident.1 = id3, ident.2 = id4, min.pct = 0.25, logfc.threshold = 1)
imm3 <- subset(imm3, p_val_adj < 0.01)

imm12 <- FindMarkers(se.dc.mus,ident.1 = id5, ident.2 = id6, min.pct = 0.25, logfc.threshold = 1)
imm12 <- subset(imm12, p_val_adj < 0.01)

cmm_g <- rownames(cmm)
imm3_g <- rownames(imm3)
imm12_g <- rownames(imm12)

infection3h_maturity_specific <- setdiff(imm3_g, cmm_g)
infectionc_specific <- setdiff(cmm_g, imm3_g)
infection12h_maturity_specific <- setdiff(imm12_g, cmm_g)
infectionc_specific2 <- setdiff(cmm_g, imm12_g)
control_maturity_specific_markers <- union(infectionc_specific,infectionc_specific2)

infection3h_maturity_specific <- setdiff(infection3h_maturity_specific, infection12h_maturity_specific)
infection12h_maturity_specific <- setdiff(infection12h_maturity_specific, infection3h_maturity_specific)

DEG_maturation_infection3h <- subset(imm3, rownames(imm3) %in% infection3h_maturity_specific)
DEG_maturation_infection12h <- subset(imm12, rownames(imm12) %in% infection12h_maturity_specific)

DEG_maturation_infection12h

mat_12h_cor <- cor(t(as.matrix(se.dc.mus@assays$SCT@data[rownames(DEG_maturation_infection12h), ])))

corrplot(mat_12h_cor, method="color", tl.col = "black", order = "FPC", col=colorRampPalette(c("blue4","white","red3"))(100), cl.lim = c(-1,1))


```

I would like to compare the differential expressed genes for LDM and PTG at 12h between mature and immature cells in more detail and display them in a nicer fashion than a volcano-plot, therefore I subset the data to only have mature and immature PTG and LDM unfected cells at 12 hours

```{r, subset object}

se.12h <- SubsetData(se.dc.mus, cells = rownames(subset(se.dc.mus[[]], cell_class_maturation %in% c("PTG_12h_mature","PTG_12h_immature", "LDM_12h_mature", "LDM_12h_immature"))))

se.12h <- SetIdent(se.12h, value = "cell_class_maturation")
se.12h.m<- FindAllMarkers(se.12h, logfc.threshold = 1, min.pct = 0.25, only.pos = T)
se.12h.m  <- subset(se.12h.m, p_val_adj < 0.01)

se.12h.m  %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)

DoHeatmap(se.12h, features = se.12h.m$gene)


```

```{r, pairwise DEG cluster and timepoint based}

#Paste the cluster to the respective cell_class of each gene and add as a new column in the metadata 
se.dc.mus$cluster_and_cell_class <- paste0(se.dc.mus$SCT_snn_res.0.8, "_", se.dc.mus$cell_class)

id1 <- c("2_LDM_12h", "2_PTG_12h")
id2 <- c("3_PTG_12h", "5_LDM_12h")
#Set identity
se.dc.mus <- SetIdent(se.dc.mus, value = "cluster_and_cell_class")
m12h <- FindMarkers(object = se.dc.mus, ident.1 = id1, ident.2 = id2, logfc.threshold = 0.5, min.pct = 0.25)
m12h<- subset(m12h, p_val_adj < 0.05)
m12h

#Only for LDM12h

id3 <- c("2_LDM_12h")
id4 <- c("5_LDM_12h")

m12hLDM <- FindMarkers(object = se.dc.mus, ident.1 = id3, ident.2 = id4, logfc.threshold = 0.5, min.pct = 0.25)
m12hLDM <- subset(m12hLDM, p_val_adj < 0.05)
m12hLDM

id5 <- c("2_PTG_12h")
id6 <- c("5_PTG_12h")

m12hPTG <- FindMarkers(object = se.dc.mus, ident.1 = id5, ident.2 = id6, logfc.threshold = 0.5, min.pct = 0.25)
m12hPTG <- subset(m12hPTG, p_val_adj < 0.05)
m12hPTG


#Make Volcano-plots to visualize the DGE 
volcano.12h.cluster <- EnhancedVolcano(m12h, lab = rownames(m12h),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.12h.LDM.cluster <- EnhancedVolcano(m12hLDM, lab = rownames(m12hLDM),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.12h.PTG.cluster <- EnhancedVolcano(m12hPTG, lab = rownames(m12hPTG),x= "avg_logFC", y = "p_val_adj", xlim = c(-6,6), title = NULL, subtitle =NULL, pCutoff = 0.01, FCcutoff =  1, legendLabSize = 6, legendPosition = "right", axisLabSize = 10)

volcano.12h.cluster
volcano.12h.LDM.cluster
volcano.12h.PTG.cluster

m12h <- subset(m12h, avg_logFC > 1 & p_val_adj < 0.01)
m12hLDM <- subset(m12hLDM, avg_logFC > 1 & p_val_adj < 0.01)
m12hPTG <- subset(m12hPTG, avg_logFC > 1 & p_val_adj < 0.01)

#add gene descriptions

genetable.12h <- subset(genes.table, external_gene_name %in% rownames(m12h))
genetable.12hLDM <- subset(genes.table, external_gene_name %in% rownames(m12hLDM))
genetable.12hPTG <- subset(genes.table, external_gene_name %in% rownames(m12hPTG))

m12h$external_gene_name <- rownames(m12h)
m12hLDM$external_gene_name <- rownames(m12hLDM)
m12hPTG$external_gene_name <- rownames(m12hPTG)

m12h <- dplyr::left_join(m12h, genetable.12h, by = "external_gene_name")
m12hLDM <- dplyr::left_join(m12hLDM, genetable.12hLDM, by = "external_gene_name")
m12hPTG <- dplyr::left_join(m12hPTG, genetable.12hPTG, by = "external_gene_name")


write.xlsx(m12h, "~/Dropbox/shared_ankarklev/Franzi-analysis/pairwise-DEG-12h-cluster2against3and5.xlsx", row.names =TRUE)
write.xlsx(m12hLDM, "~/Dropbox/shared_ankarklev/Franzi-analysis/pairwise-DEG-12h-cluster2against5.xlsx", row.names =TRUE)
write.xlsx(m12hPTG, "~/Dropbox/shared_ankarklev/Franzi-analysis/pairwise-DEG-12h-cluster2against3.xlsx", row.names =TRUE)

png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-12h-cluster2-cluster3and5.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.12h.cluster)
dev.off()

png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-12h-LDM-cluster2-cluster5.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.12h.LDM.cluster)
dev.off()

png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/volcanoplot-12h-PTG-cluster2-cluster3.png", width = 15, height = 25, units = "cm", res = 300)
print(volcano.12h.PTG.cluster)
dev.off()

```


Integrate toxodata in the mouse data 


```{r, add toxoclusters to metadata and highlight them in umap}

toxo_clusters <- as.data.frame(paste0("toxo", "_", se.toxo$SCT_snn_res.0.8))
rownames(toxo_clusters) <- rownames(se.toxo[[]])
names(toxo_clusters)[1] <- "toxo_cluster"
head(toxo_clusters)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_clusters, col.name = "toxoclusters")

#Visulize only cells of toxo_cluster = toxo_2 on DC uMAP!

se.dc.mus <- SetIdent(se.dc.mus, value = "toxoclusters")

toxo.dc.0 <- DimPlot(se.dc.mus, cols = c("darkred", "gray48", "gray48", "gray48", "gray48", "gray48", "gray48"))


toxo.dc.1 <- DimPlot(se.dc.mus, cols = c("gray48", "darkred", "gray48", "gray48", "gray48", "gray48", "gray48"))

toxo.dc.2 <- DimPlot(se.dc.mus, cols = c("gray48", "gray48","darkred" ,"gray48", "gray48", "gray48", "gray48"))
  
toxo.dc.3 <- DimPlot(se.dc.mus, cols = c("gray48",  "gray48", "gray48", "darkred","gray48", "gray48", "gray48"))

toxo.dc.4 <- DimPlot(se.dc.mus, cols = c("gray48",  "gray48", "gray48","gray48",  "darkred","gray48", "gray48"))

png("~/Dropbox/shared_ankarklev/Franzi-analysis/Results-FiguresAndTables/Toxo-Clusters-DC-Umap.png", width = 40, height = 25, unit= "cm", res= 300)
print(cowplot::plot_grid(toxo.dc.0, toxo.dc.1, toxo.dc.2, toxo.dc.3, toxo.dc.4))
dev.off()
  
```

```{r, DEG toxoclusters in DC data}
#Set identity to Toxoclusters 

se.dc.mus <- SetIdent(se.dc.mus, value = "toxoclusters")

toxocluster.dc.markers <- FindAllMarkers(se.dc.mus, logfc.threshold = 1, min.pct = 0.25, only.pos = T)
toxocluster.dc.markers  <- subset(toxocluster.dc.markers, p_val_adj < 0.01)

toxocluster.dc.markers  %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)

se.dc.mus <- SetIdent(se.dc.mus, value = "seurat_clusters")
toxocluster.dc.heatmap <- DoHeatmap(se.dc.mus, features = toxocluster.dc.markers$gene)

toxocluster.dc.heatmap

#Add gene descriptions: 
genetabletoxoclusters <- subset(genes.table, external_gene_name %in% rownames(toxocluster.dc.markers))
toxocluster.dc.markers$external_gene_name <- rownames(toxocluster.dc.markers)


toxocluster.dc.markers<- dplyr::left_join(toxocluster.dc.markers, genetabletoxoclusters, by = "external_gene_name")


write.xlsx(toxocluster.dc.markers, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-spreadsheets/Toxoclusters-DC-Markers.xlsx")
```


```{r, pairwise DEG toxo-cluster and maturation}

#Paste the cluster to the respective cell_class of each gene and add as a new column in the metadata 
se.dc.mus$toxocluster_maturation <- paste0(se.dc.mus$toxoclusters, "_", se.dc.mus$maturation_state)

id0 <- "toxo_0_immature"
id10 <- "toxo_0_mature"
#Set idenity
se.dc.mus <- SetIdent(se.dc.mus, value = "toxocluster_maturation")
toxo_dc_c0 <- FindMarkers(object = se.dc.mus, ident.1 = id0, ident.2 = id10, logfc.threshold = 2)
toxo_dc_c0 <- subset(toxo_dc_c0, p_val_adj < 0.01)
#Set filters! 
toxo_dc_c0

id1 <- "toxo_1_immature"
id2 <- "toxo_1_mature"
#Set idenity
se.dc.mus <- SetIdent(se.dc.mus, value = "toxocluster_maturation")
toxo_dc_c1 <- FindMarkers(object = se.dc.mus, ident.1 = id1, ident.2 = id2, logfc.threshold = 2)
toxo_dc_c1 <- subset(toxo_dc_c1, p_val_adj < 0.01)
#Set filters! 
toxo_dc_c1

id3 <- "toxo_2_immature"
id4 <- "toxo_2_mature"
#Set idenity
se.dc.mus <- SetIdent(se.dc.mus, value = "toxocluster_maturation")
toxo_dc_c2 <- FindMarkers(object = se.dc.mus, ident.1 = id3, ident.2 = id4, logfc.threshold = 2)
toxo_dc_c2 <- subset(toxo_dc_c2, p_val_adj < 0.01)
#Set filters! 
toxo_dc_c2

id5 <- "toxo_3_immature"
id6 <- "toxo_3_mature"
#Set idenity
se.dc.mus <- SetIdent(se.dc.mus, value = "toxocluster_maturation")
toxo_dc_c3 <- FindMarkers(object = se.dc.mus, ident.1 = id5, ident.2 = id6, logfc.threshold = 2)
toxo_dc_c3 <- subset(toxo_dc_c3, p_val_adj < 0.01)
#Set filters! 
toxo_dc_c3


#Forth cluster has to little cells to run DGEA!!
id7 <- "toxo_4_immature"
id8 <- "toxo_4_mature"
#Set idenity
se.dc.mus <- SetIdent(se.dc.mus, value = "toxocluster_maturation")
toxo_dc_c4 <- FindMarkers(object = se.dc.mus, ident.1 = id7, ident.2 = id8, logfc.threshold = 2)
toxo_dc_c4 <- subset(toxo_dc_c4, p_val_adj < 0.01)
#Set filters! 
toxo_dc_c4


##Add gene descriptions:

genetableC0 <- subset(genes.table, external_gene_name %in% rownames(toxo_dc_c0))
genetableC1 <- subset(genes.table, external_gene_name %in% rownames(toxo_dc_c1))
genetableC2 <- subset(genes.table, external_gene_name %in% rownames(toxo_dc_c2))
genetableC3 <- subset(genes.table, external_gene_name %in% rownames(toxo_dc_c3))


toxo_dc_c0$external_gene_name <- rownames(toxo_dc_c0)
toxo_dc_c1$external_gene_name <- rownames(toxo_dc_c1)
toxo_dc_c2$external_gene_name <- rownames(toxo_dc_c2)
toxo_dc_c3$external_gene_name <- rownames(toxo_dc_c3)

toxo_dc_c0 <- dplyr::left_join(toxo_dc_c0, genetableC0, by = "external_gene_name")
toxo_dc_c1 <- dplyr::left_join(toxo_dc_c1, genetableC0, by = "external_gene_name")
toxo_dc_c2 <- dplyr::left_join(toxo_dc_c2, genetableC0, by = "external_gene_name")
toxo_dc_c3<- dplyr::left_join(toxo_dc_c3, genetableC0, by = "external_gene_name")



write.xlsx(toxo_dc_c0, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-spreadsheets/ToxoCluster0-DC-immature(negative)-mature(positive.xlsx")
write.xlsx(toxo_dc_c1, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-spreadsheets/ToxoCluster1-DC-immature(negative)-mature(positive.xlsx")
write.xlsx(toxo_dc_c2, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-spreadsheets/ToxoCluster2-DC-immature(negative)-mature(positive.xlsx")
write.xlsx(toxo_dc_c3, "~/Dropbox/shared_ankarklev/Franzi-analysis/DEG-spreadsheets/ToxoCluster3-DC-immature(negative)-mature(positive.xlsx")

##Get numbers of cells for each cluster

```

```{r, visualize the cells that are Gra15+ and Rop16+ in the DC set}

toxo_gra <- as.data.frame(paste0(se.toxo$Gra15_expression))
rownames(toxo_gra) <- rownames(se.toxo[[]])
names(toxo_gra)[1] <- "gra-expression"
head(toxo_gra)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_gra, col.name = "Gra15_expression")


toxo_rop <- as.data.frame(paste0(se.toxo$Rop16_expression))
rownames(toxo_rop) <- rownames(se.toxo[[]])
names(toxo_rop)[1] <- "rop-expression"
head(toxo_rop)

se.dc.mus <- AddMetaData(se.dc.mus, metadata = toxo_rop, col.name = "Rop16_expression")

DimPlot(se.dc.mus, group.by = "Rop16_expression")

se.dc.rop <- subset(se.dc.mus, subset = Rop16_expression %in% "Rop16+", slot = "data")
se.dc.gra <- subset(se.dc.mus, subset = Gra15_expression %in% "Gra15+", slot = "data")
DimPlot(se.dc.rop, group.by = "toxoclusters")
DimPlot(se.dc.gra, group.by = "cell_class")

DimPlot(se.dc.mus, group.by = "Gra15_expression")
DimPlot(se.dc.mus, group.by = "Rop16_expression")

```

